### 2016-04-13 Create_Summary_Tables_AllFishingActivities.Rmd
Added files via upload

---
title: "Permits, permit holders, vessels 2"
author: "Jordan Watson"
date: "March 9, 2016"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document: default
---

```{r chunkSet, cache=FALSE, echo=FALSE,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(knitr)
library(reshape2)
library(vegan)
#setwd("~/Desktop/NCEAS")
setwd("//nmfs.local/AKC-ABL/Users2/jordan.watson/Desktop/AFSC/GOA/AKFIN/Meeting")

opts_chunk$set(tidy=TRUE, message=FALSE)
```

```{r loadData,warning=FALSE,message=FALSE}

load("Data/cfecCleaned_new.RData")

# initially only use data 1985+
cfec = cfec[cfec$year > 1984, ]
cfec <- rename(cfec,adfg=cadfg) %>% data.frame
cfec$year <- as.numeric(cfec$year)

#  Read-in the CPI adjusted for 2009 base dollars. 
cpi <- read.csv("Data/CPI_2009.csv") %>% filter(year>1984)
cfec <- left_join(cfec,cpi) 
#  Adjust g_earn and g_price to 2009 dollars
cfec[,c("g_earn","g_price")] <- cfec[,c("g_earn","g_price")] / cfec$GDPDEF09_base1
cfec$GDPDEF09_base1 <- NULL
rm(cpi)
```

```{r}
#  Get rid of the spaces in p_fshy
cfec$p_fshy <- gsub(" ","",cfec$p_fshy)

#  Read in the fishing locations by fishery permit (p_fshy)
p_fshy <- read.csv("Data/new.locs.csv")
p_fshy$p_fshy <- as.character(p_fshy$p_fshy)

#  Combine these areas (original and our coarse new area) to the data
cfec <- left_join(cfec,p_fshy)

#  Now for gears...
#  Read in gear description file and link with cfec.
gears <- read.csv("Data/new_gear_cat.csv")
cfec <- left_join(cfec,gears[,c("gearn","Gear")])

#  Troll fisheries only happen in Southeast
cfec$New.Loc[cfec$New.Loc=="Statewide" & cfec$Gear=="Troll"] <- "Southeast"
#  Abalone were only ever caught in southeast
cfec$New.Loc[cfec$New.Loc=="Statewide" & cfec$spec=="ABLN"] <- "Southeast"
```

```{r}
#  Create a temporary (ie., smaller) data frame for calculating with only the necessary fields.
temp2 <- select(cfec,-F_REGION,-F_INPFC,-g_price,-P_TYPE,-P_STATUS,-A_RES,-g_price,-startdt,-landdate,harvest,F_PERMIT)  %>% filter(g_earn>0 & !is.na(Gear) & !is.na(New.Loc))

#  Identify the earliest fishing year for each p_holder
startyear <- group_by(temp2,p_holder) %>% summarise(start=min(year))
#  Join that year to the data
temp2 <- left_join(temp2,startyear)
rm(startyear)

#  Define custom function for calculating the mode
Mode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
}

#  Calculate initial metrics for each p_holder and year
xx <- group_by(temp2,p_holder,year) %>% 
              summarise(StartYear=start[1],
              permits=length(unique(p_serial)),# Number of permits
              permitTypes=length(unique(p_fshy)),# Number of permit types
              landings=length(unique(f_ticket)), # Number of fish tickets (proxy for # of trips)
              boats=length(unique(adfg)),# Number of boats fished by p_holder
              areaSm=length(unique(Location)), # Number of permit areas held (N=51)
              areaLg=length(unique(New.Loc)), # Number of our aggregated permit areas held (N=13)
              statareas=length(unique(stat6)), # Number of different areas from which fish were reported to have been caught
              vlength=Mode(VLength)
              ) %>% data.frame

rm(gears,p_fshy,Mode)
```

### We now begin to examine Revenues as a function of fishing area and gear for each p_holder and year

#### Aggregate revenue by area and gear for each p_holder and year
```{r}
#temp.dollars <- group_by(temp2,p_holder,year,Gear,New.Loc) %>% summarise(Dollars=sum(g_earn))
#
#for(i in 1:length(unique(temp.dollars$Gear))){     
#  print(ggplot(temp.dollars[temp.dollars$Gear==paste(unique(temp.dollars$Gear)[i]),],aes(x=factor(year),y=log(Dollars))) + geom_boxplot() + facet_wrap(~New.Loc) + ggtitle(paste(unique(temp.dollars$Gear)[i])))
#}
```

```{r}
#  Tally the dollars of fish by each p_holder during each year for each gear type
#  First in long format
tempgear.d <- group_by(temp2,p_holder,year,Gear) %>% summarise(Dollars=sum(g_earn))
#  Now in wide format, with zeros filled in.
tempgearwide.d <- dcast(tempgear.d,p_holder+year~Gear,value.var="Dollars")
tempgearwide.d[is.na(tempgearwide.d)] <- 0

#  Tally the dollars of fish by each p_holder during each year for each permit location
#  First in long format
temparea.d <- group_by(temp2,p_holder,year,New.Loc) %>% summarise(Dollars=sum(g_earn))
#  Now in wide format, with zeros filled in.
tempareawide.d <- dcast(temparea.d,p_holder+year~New.Loc,value.var="Dollars")
tempareawide.d[is.na(tempareawide.d)] <- 0

#  Tally the dollars of fish by each p_holder during each year for each species
#  First in long format
tempspec.d <- group_by(temp2,p_holder,year,spec) %>% summarise(Dollars=sum(g_earn))
tempspecwide.d <- dcast(tempspec.d,p_holder+year~spec,value.var="Dollars")
#  Now in wide format
tempspecwide.d[is.na(tempspecwide.d)] <- 0

# Join each of the above wide format tables. In joining, values of g_earn from different tables will become redundant so do not use these dollar values.   
tempjoin.dollars <- inner_join(xx,tempareawide.d) %>% inner_join(tempgearwide.d) %>% inner_join(tempspecwide.d)
# Convert the dollar values to dummy values
tempjoin.dollars[,-c(which(names(tempjoin.dollars) %in% colnames(xx)))] <- lapply(tempjoin.dollars[,-c(which(names(tempjoin.dollars) %in% colnames(xx)))], function(x) replace(x,x!=0, 1))
# Join the total annual revenue for each p_holder with the above table
revenueMetrics <- group_by(temp2,p_holder,year) %>% summarise(totalRev=round(sum(g_earn)),
                                                              medianRev=round(median(g_earn)),
                                                              cvRev=round(sd(g_earn)/mean(g_earn)))  %>% 
                                                              inner_join(tempjoin.dollars) %>% 
                                                              mutate(yearsFishing=(year-StartYear)+1) %>%
                                                              select(p_holder,year,yearsFishing,everything()) %>% data.frame

# For plotting purposes the long format will be more useful
longjoin.dollars <- inner_join(xx,temparea.d) %>% inner_join(tempgear.d) %>% inner_join(tempspec.d) %>% mutate(yearsFishing=(year-StartYear)+1)
```

Now do the same thing but aggregate the pounds landed (instead of the revenue) landed by area, gear,species for each p_holder and year
```{r}

#  Tally the pounds of fish by each p_holder during each year for each gear type
#  First in long format
tempgear.p <- group_by(temp2,p_holder,year,Gear) %>% summarise(Mass=sum(g_pounds))
#  Now in wide format, with zeros filled in.
tempgearwide.p <- dcast(tempgear.p,p_holder+year~Gear,value.var="Mass")
tempgearwide.p[is.na(tempgearwide.p)] <- 0

#  Tally the pounds of fish by each p_holder during each year for each permit location
#  First in long format
temparea.p <- group_by(temp2,p_holder,year,New.Loc) %>% summarise(Mass=sum(g_pounds))
#  Now in wide format, with zeros filled in.
tempareawide.p <- dcast(temparea.p,p_holder+year~New.Loc,value.var="Mass")
tempareawide.p[is.na(tempareawide.p)] <- 0

#  Tally the pounds of fish by each p_holder during each year for each species
#  First in long format
tempspec.p <- group_by(temp2,p_holder,year,spec) %>% summarise(Mass=sum(g_pounds))
tempspecwide.p <- dcast(tempspec.p,p_holder+year~spec,value.var="Mass")
#  Now in wide format
tempspecwide.p[is.na(tempspecwide.p)] <- 0

# For plotting purposes the long format will be more useful
longjoin.pounds <- inner_join(xx,temparea.p) %>% inner_join(tempgear.p) %>% inner_join(tempspec.p) %>% mutate(yearsFishing=(year-StartYear)+1)

tempjoin.pounds <- inner_join(xx,tempareawide.p) %>% inner_join(tempgearwide.p) %>% inner_join(tempspecwide.p)
tempjoin.pounds[,-c(which(names(tempjoin.pounds) %in% colnames(xx)))] <- lapply(tempjoin.pounds[,-c(which(names(tempjoin.pounds) %in% colnames(xx)))], function(x) replace(x,x!=0, 1))

# Join the total annual revenue for each p_holder with the above table
landedMetrics <- group_by(temp2,p_holder,year) %>% summarise(totalPounds=round(sum(g_pounds)),
                                                              medianPound=round(median(g_pounds)),
                                                              cvPounds=round(sd(g_pounds)/mean(g_pounds)))  %>% 
                                                              inner_join(tempjoin.pounds) %>% 
                                                              mutate(yearsFishing=(year-StartYear)+1) %>%
                                                              select(p_holder,year,yearsFishing,everything()) %>% data.frame

```

Combine metrics into a single table

```{r}

AllMetrics.wide <- inner_join(revenueMetrics,landedMetrics) %>% select(p_holder,StartYear,yearsFishing,vlength,permits:statareas,totalRev:cvRev,totalPounds:cvPounds,everything())
AllMetrics.long <- inner_join(longjoin.pounds,longjoin.dollars) %>% select(p_holder,StartYear,yearsFishing,vlength,permits:New.Loc,Gear,spec,Mass,Dollars)

temparea <- inner_join(temparea.p,temparea.d)
tempgear <- inner_join(tempgear.p,tempgear.d)
tempspec <- inner_join(tempspec.p,tempspec.d)

rm(xx,cfec,temp2,revenueMetrics,landedMetrics,longjoin.pounds,longjoin.dollars,tempjoin.dollars,tempjoin.pounds,tempspec.p,tempspec.d,temparea.p,temparea.d,tempgear.p,tempgear.d)
save.image("cfecModelsetup.RData")
```

## Metadata - what do all of the different objects mean?
For modeling, the object most usefull will likely be "AllMetrics.wide"

*AllMetrics.long* contains combined characteristics for each p_holder and a single column for all species, a single column for all gears, a single column for all areas (New.Loc), a single column for pounds landed, and a single column for dollars earned.

*AllMetrics.wide* contains combined characteristics for each p_holder and a different column for each species, a different column for each gear, a different column for each areas (New.Loc), a single column total,median and cv of revenue, a single column for total, median, and cv of pounds landed.

*temparea* contain dollars & pounds for each p_holder by year and by area (New.Loc) - all areas are in a single column

*tempareawide.d* / *tempareawide.p* contain dollars / pounds for each p_holder by year and by area (New.Loc) - each areas is in a different column

*tempgear* contain dollars & pounds for each p_holder by year and by gear type - all gears are in a single column

*tempgearwide.d* / *tempgearwide.p* contain dollars / pounds for each p_holder by year and by gear - each gear is in a different column

*tempspec* contain dollars & pounds for each p_holder by year and by species - all species are in a single column

*tempspecwide.d* / *tempspecwide.p* contain dollars / pounds for each p_holder by year and by species - each species is in a different column