### 2016-06-14 portfolio/analysis/1.5-createData_linear_modeling.Rmd

---
title: "createData_linear_modeling"
author: ""
date: "June 05, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
library(metafolio)
library(randomForest)
library(date)

#load("cfecCleaned2.Rdata")
cfec <- feather::read_feather("data/cfec.feather")

# convert 6-dig p_fshy codes to 5 digits
# removing blank space - now they're all length(5)
idx = which(nchar(cfec$p_fshy)==6)
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1), 
  substr(cfec$p_fshy[idx], 3, 6), sep="")

# remove white space in 2nd position
idx = which(substr(cfec$p_fshy, 2, 2) == " ")
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1), 
  substr(cfec$p_fshy[idx], 3, length(cfec$p_fshy[idx])), sep="")

region<-read.csv("areas/regions.csv")
region<-region[,c(1:2)] #don’t include “assigned by” column
cfec<-left_join(cfec,region)
#cfec$region <- replace(cfec$region, is.na(cfec$region), "Unassigned")
```

We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).

https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0

```{r, echo=FALSE}
deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
0.91981, 0.94812, 0.97334, 0.9925, 1,  
1.01217, 1.03307, 1.05213, 1.06926, 1.08682))

# Adjust price and g_earn for inflation
cfec$year = as.numeric(cfec$year)
cfec$day = substr(cfec$landdate, 6, 10)
cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
```

### Function definiton  
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity. Note the C++ version is commented out below,

```{r, warning = FALSE, message=FALSE, echo=FALSE}

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(p_holder, year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear2 = function(dataFrame, variable1, variable2) {
  effDiv = group_by_(dataFrame, variable1, variable2, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>% 
  group_by_("p_holder", "year", variable2) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

effectiveDiversity_by_Year = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity = function(dataFrame) {
  effDiv = group_by(dataFrame, year) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity2 = function(dataFrame, variable2) {
  effDiv = group_by_(dataFrame, "year", variable2) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}
```

## Filtering out unique vessel - person combinations  
Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year. Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.

Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.

```{r}
cfecnew = cfec
cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
  summarize(nBoat = length(unique(cadfg))) %>%
  filter(nBoat==1) %>% 
  left_join(cfecnew)

# Deal with setnet fisheries, where lots of p_holder
# share same codes
cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
  summarize(nPerson = length(unique(p_holder))) %>%
  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
  left_join(cfec.1boat)

cfecnew = cfec.1boat
dim(cfecnew)[1]
```

Aggregate annual metrics by person,

```{r}

cfecnew$day = mdy.date(as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 1)), as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 2)), cfecnew$year) - mdy.date(1, 1, cfecnew$year)

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg2[1], 
    days = diff(range(day) + 1), 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    length = mean(VLength, na.rm=T))

cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
  filter(npermit == 1)

# also calculate weighted simpson diversity of gears and species
cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, specn) %>% 
  summarize(totRev.spec = sum(g_earn)) %>% 
  group_by(p_holder, year) %>% 
  summarize(specDiv = simp.div(totRev.spec)))

# also calculate weighted simpson diversity of gears and species
cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, gearn) %>% 
  summarize(totRev.gear = sum(g_earn)) %>% 
  group_by(p_holder, year) %>% 
  summarize(gearDiv = simp.div(totRev.gear)))

#temp = group_by(cfecnew, p_holder, year, specn) %>% 
#  summarize(totRev.spec = sum(g_earn), 
#    permit = p_fshy[1])
#temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
#    summarize(totRev = sum(g_earn)))
#temp$totRev.spec = temp$totRev.spec / temp$totRev

#mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
#  summarize(totRev.mkt.spec = sum(g_earn)) 
#mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
#  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
#mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
#names(mkt)[1] = "permit"

# combine individual and market datasets
#temp = left_join(temp, mkt)

# Calculate squared proportions
#temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2

#cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))

# Also merge in previous revenue. Idx stores index of people last year
idx = match(paste(cfecAnnual$p_holder,cfecAnnual$year-1), paste(cfecAnnual$p_holder,cfecAnnual$year))

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
#cfecAnnual$p2.prev = cfecAnnual$p2[idx]

cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev

# Filter out salmon people
cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", "Y", "N")

cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]

cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]

saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
saveRDS(cfecAnnual.nonsalmon, file="salmon_linearModeling_complete.rds")
```

We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.

```{r}

cfecnew$area = substr(cfecnew$p_fshy,4,4)
cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
cfecnew$area[cfecnew$area == "B"] = "Statewide"
cfecnew$area[cfecnew$area == "D"] = "Yakutat"
cfecnew$area[cfecnew$area == "E"] = "PWS"
cfecnew$area[cfecnew$area == "H"] = "CookInlet"
cfecnew$area[cfecnew$area == "J"] = "Westward"
cfecnew$area[cfecnew$area == "K"] = "Kodiak"
cfecnew$area[cfecnew$area == "L"] = "Chignik"
cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
cfecnew$area[cfecnew$area == "N"] = "Nelson"
cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
cfecnew$area[cfecnew$area == "T"] = "BristolBay"
cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
cfecnew$area[cfecnew$area == "Z"] = "NortonSound"

cfecnew$spec = substr(cfecnew$p_fshy,1,1)
cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")

cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
  summarize(npermit = length(unique(p_fshy)), 
    allpermits = paste(sort(unique(p_fshy)), collapse = " "),
    areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
  filter(npermit > 1)

```

## Model fitting

Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).

```{r}
library(lme4)

cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)

full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
  as.data.frame %>%
  group_by(permit) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% 
  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
```