### 2016-06-15 portfolio/analysis/4b-Bayesian_portfolio_modeling.Rmd
fixes
I modified the code to allow the 4 arrays of ints to still be an array
with 1 random effect. Previously, a dimension error was being thrown by
STAN.

---
title: "Bayesian_portfolio_modeling"
author: ""
date: "June 05, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load packages and such 
```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
library(metafolio)
library(randomForest)
library(date)
library(rstan)

```

## LMER Model fitting

Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).

```{r}
cfecAnnual.salmon = readRDS(file="salmon_linearModeling_complete.rds")

cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)

full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
  as.data.frame %>%
  group_by(permit) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% 
  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
```

## Code for testing fitting the model using STAN.

```{r}
cfecAnnual.nonsalmon = readRDS(file="nonsalmon_linearModeling_complete.rds")

# Subset the data to include only p_holder - year combinations with > 100 records, basically the top 20 permits
testData = cfecAnnual.nonsalmon[cfecAnnual.nonsalmon$permit%in%names(rev(sort(table(cfecAnnual.nonsalmon$permit)))[1:20]),]

dim(testData)

# Create numeric representation of characters/factors for linear modeling
P_HOLDER = as.numeric(as.factor(testData$p_holder))
PERMIT = as.numeric(as.factor(testData$permit))
PERMIT_AREA = as.numeric(as.factor(substr(testData$permit, 4, 4)))
PERMIT_SPEC = as.numeric(as.factor(substr(testData$permit, 1, 1)))
SPEC_DIV = testData$specDiv
DAYS = log(testData$days+1)

# Calculate variables / matrices to plug into STAN. These variables / matrices can 
# be changed to add or delete covariates. Covariates can be included in the following:
# fixed effects in slope of mean (b1_cov)
# fixed effects in slope of var (sigma1_cov)
# random effects of intercept of mean (b0_re)
# random effects of intercept of var (sigma0_re)
# random effects of slope of mean (b1_re)
# random effects of slope of var (sigma1_re)

N = dim(testData)[1]
b1_cov = cbind(SPEC_DIV) # fixed effect covariate on slope
sigma1_cov = cbind(SPEC_DIV) # fixed effect covariate on slope
b1_cov_re = cbind(SPEC_DIV)
sigma1_cov_re = cbind(SPEC_DIV)

b0_re = cbind(PERMIT) #cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
b0_re_levels = apply(b0_re,2,max)
b0_re_n = dim(b0_re)[2]

b1_re = cbind(PERMIT)#cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
b1_re_levels = apply(b1_re,2,max)
b1_re_n = dim(b1_re)[2]

sigma0_re = cbind(PERMIT)
sigma0_re_levels = apply(sigma0_re,2,max)
sigma0_re_n = dim(sigma0_re)[2]

sigma1_re = cbind(PERMIT)
sigma1_re_levels = apply(sigma1_re,2,max)
sigma1_re_n = dim(sigma1_re)[2]

K_mean = dim(b1_cov)[2]
K_sig = dim(sigma1_cov)[2]
K_mean_re = dim(b1_cov_re)[2]
K_sig_re = dim(sigma1_cov_re)[2]
y = log(testData$revenue)

### Center all of the covariate matrices
#b1_cov = scale(b1_cov, scale=FALSE)
#sigma1_cov = scale(sigma1_cov, scale=FALSE)
#b1_cov_re = scale(b1_cov_re, scale=FALSE)
#sigma1_cov_re = scale(sigma1_cov_re, scale=FALSE)
```

```{r}
# Named list of data
list_data = list("N"=N, "K_mean"=K_mean, "b1_cov"=b1_cov, "sigma1_cov"=sigma1_cov,
  "K_mean_re"=K_mean_re, "K_sig_re"=K_sig_re,"b1_cov_re"=b1_cov_re, "sigma1_cov_re"=sigma1_cov_re,
  "b0_re"=b0_re, "b0_re_levels"=c(b0_re_levels,0), "b0_re_n"=b0_re_n,
  "b1_re"=b1_re, "b1_re_levels"=c(b1_re_levels,0), "b1_re_n"=b1_re_n,
  "sigma0_re"=sigma0_re, "sigma0_re_levels"=c(sigma0_re_levels,0), "sigma0_re_n"=sigma0_re_n,
  "sigma1_re"=sigma1_re, "sigma1_re_levels"=c(sigma1_re_levels,0), "sigma1_re_n"=sigma1_re_n, "K_sig"=K_sig, "y"=y)
# parameters to monitor
list_pars = c("b0_mu","b0_dev","b0_tau", "b1_mu","b1_dev","b1_tau", "sigma0_mu","sigma0_dev","sigma0_tau", "sigma1_mu","sigma1_dev","sigma1_tau")

# fit the model
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

stanMod = stan(file = 'portfolio_mixedModel.stan',data = list_data, 
  verbose = TRUE, chains = 1, thin = 1, warmup = 20, iter = 50, 
  pars = list_pars)
p = extract(stanMod, list_pars, permuted=TRUE)
```

## To-do list:
### 1. model covariance of slope / intercept in random effects. maybe start with bivariate normal
### 2. Make plots of random effects v permits