### 2016-06-29 portfolio/text/supp.tex
Add Eric's notes on the differenced model

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{amsmath}
\usepackage{fixltx2e}

\begin{document}

\section{Log-difference model}

The log-difference model has several pros and cons over the undifferenced model. Negatives are that it uses a smaller subset of the data; positives are that it doesn't require estimating people-specific random effects (n = 3000+) and removes possible time effects. The basic regression model is

\begin{equation}
  { Y }_{ i,t } \sim \mathrm{N}(\hat { { Y }}_{ i,t } ,\sigma^2 )
\end{equation}

\subsection{Model for the mean}

The model's expectation is written as a hierarchical random effects linear model with random effects in the intercept and slope.

\begin{equation}
{ \hat { Y }  }_{ t,i }={ B }_{ 0,j }+{ B }_{ 1,j }\cdot { s }_{ t,i }+{o}_{ t,i}
\end{equation}

where ${ B }_{ 0,j }$ is a normally distributed random effect intercept in the mean (j subscripting different strategies). The interpretation of ${B}_{ 0,j}$ is that each after log-differencing, each strategy has different effects on changes in revenue (some tend to be winners, others tend to be losers). Initial versions of this model included estimating ${ B }_{ 0,j }$, which were largely centered on 0 (as expected), so these random effects were set to 0.  
\newline

The covariate ${ s }_{ t,i }$ is input as data, and represents the percent change in species diversity over the same time period that the percent change in revenue is being modeled, $log\left( \frac { { div }_{ t+1,i } }{ { div }_{ t,i } }  \right)$ where ${ div }_{ t,i }$ is the effective diversity of individual *i* at time *t*. The coefficient ${ B }_{ 1,j }$ is a random effect on the slope, by strategy, representing how much an X% change in species diversity within a strategy translates to a Y% change in revenue.

\subsection{Offset}

To account for variable effort over time (changes in revenue could be attributed to diversification, but also to more fishing) we included an offset term ${o}_{t,i}$, representing the percent change in landing dates between years. We experminted with different versions of the offset and when included in the model as a predictor, the coefficient is estimated to be ~ 1, so this is a reasonable assumption. The offset is calculated as $log\left( \frac { { day }_{ t+1,i } }{ { day }_{ t,i } }  \right)$. 
\newline

\subsection{Equation for the variance}

Finally, the residual error term is also modeled as a linear function of covariates,

${ \sigma  }_{ t,i }=\sqrt { \left( \gamma _{ 0,j }+{ \gamma  }_{ 1,j }\cdot { s }_{ t,i } \right)  }$

where the $gamma _{ 0,j }$ and $gamma _{ 1,j }$ represent random intercepts and slopes in the strategies. I had been interpreting the intercepts as the between-group variability in strategies (some strategies will have more variable % change in revenue than others, similar to some stocks having higher variability in returns). I was interpreting the random slope $gamma _{ 1,j }$ as the effect of within - strategy diversification; in other words, the effect of increases in species diversification varied among different strategies.

\end{document}