### 2016-07-27 portfolio/figs/portfolio-risk-return.tex
Add old TeX file

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{graphicx}

\author{Sean Anderson}
\title{Risk returns plots broken out by dimension}
% \date{}

\begin{document}
\maketitle

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-cv.pdf}
\caption{The original plot from  before.}
\label{portfolio-gross-earnings-cv.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-semideviation.pdf}
\caption{This version shows downside semideviation.}
\label{portfolio-semideviation-gross-earnings.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-semideviation-upside.pdf}
\caption{This version shows only upside semideviation.}
\label{portfolio-gross-earnings-semideviation-upside.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6.5in]{portfolio-gross-earnings-cv-vessel-length.pdf}
\caption{These are broken out by mean vessel length as shown in the header of
each box. The panel NA represents permit holders that are associated with
multiple vessel lengths.}
\label{portfolio-gross-earnings-cv-vessel-length.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-cv-pollock.pdf}
\caption{The panel
marked TRUE should be permit holders that ever finished pollock.}
\label{port}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6.5in]{portfolio-gross-earnings-cv-taxa.pdf}
\caption{Split up in broad taxonomic categories. NA represents permit holders
that don't fall into just one of these categories.}
\label{portfolio-gross-earnings-cv-taxa.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6.5in]{portfolio-gross-earnings-cv-time-window.pdf}
\caption{The original plot broken out into a ten year
moving window. The year indicated in each panel represents the starting year.}
\label{portfolio-gross-earnings-cv-time-window.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-cv-salmon.pdf}
\caption{The right panel shows salmon permit holders and the left panel shows
everything else.}
\label{portfolio-gross-earnings-cv-salmon.pdf}
\end{center}
\end{figure}

\end{document}