### 2016-08-13 portfolio/figs/strategy-table.tex
Add strategy table to SI

% latex table generated in R 3.3.1 by xtable 1.8-2 package
Label & Description \\ 
  \midrule
Dun C & Dungeness crab \\ 
  Fin - jig & Finfish, mechanical jig \\ 
  Fin - ll & Finfish, longline, statewide \\ 
  Fin - otter & Finfish, otter trawl, statewide \\ 
  Fin - pot & Finfish, pot gear, statewide \\ 
  Hal & Halibut \\ 
  Herr - pound & Herring, pound \\ 
  Herr roe - gill & Herring roe, gillnet \\ 
  Herr roe - seine & Herring roe, purse seine \\ 
  King C & King crab \\ 
  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gill net, CI, AP, BB \\ 
  S - drift - PWS & Salmon, drift gill net, PWS \\ 
  S - drift - SE & Salmon, drift gill net, Southeast Alaska \\ 
  S - seine - AP & Salmon, purse seine, Alaska Penninsula \\ 
  S - seine - CH & Salmon, purse seine, Chignik \\ 
  S - seine - K & Salmon, purse seine, Cook Inlet and Kodiak \\ 
  S - seine - PWS & Salmon, purse seine, PWS \\ 
  S - seine - SE & Salmon, purse seine, Southeast Alaska \\ 
  S - set - NS/K & Salmon, set gillnet, Norton Sound and Kuskokwim \\ 
  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB \\ 
  S - set - YAK & Salmon, set gillnet, Yakutat \\ 
  S - set - YUK & Salmon, gillnet, lower Yukon \\ 
  S - troll & Salmon , power troll, statewide \\ 
  Sab & Sablefish, longline, statewide \\ 
  Sab - SE & Sablefish , longline, Southeast Alaska \\ 
  Sea cuc & Sea cucumber \\ 
  Shrimp & Shrimp \\ 
  Tan C & Tanner crab, pot gear vessel under 60' \\ 
  Tan C - large & Tanner crab, pot gear vessel over 60' \\ 
   \bottomrule