### 2016-08-16 portfolio/analysis/1-create-modelling-data.Rmd
Keep track of sample sizes during filtering

---
title: "Create data sets for modeling"
output:
  pdf_document:
    fig_caption: yes
---

## Differenced vs. undifferenced datasets:

1. Excluding observations that we don't have consecutive years for

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data

```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(date)

cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")

# convert 6-dig p_fshy codes to 5 digits
# removing blank space - now they're all length(5)
idx = which(nchar(cfec$p_fshy)==6)
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1),
  substr(cfec$p_fshy[idx], 3, 6), sep="")

# remove white space in 2nd position
idx = which(substr(cfec$p_fshy, 2, 2) == " ")
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1),
  substr(cfec$p_fshy[idx], 3, length(cfec$p_fshy[idx])), sep="")
```

## Function definition

We need to define a few functions for calculating diversity. We'll use the
Simpson's diversity index (HHI) to calculate the entropy-based inverse, which
is a proxy for effective species diversity.

```{r, warning = FALSE, message=FALSE, echo=FALSE}
simp.div = function(x) {
  1/sum((x/sum(x))^2)
}
```

## Previously, we excluded people using 2 boats in a year, but now we'll include them.

```{r}
# Now we'll filter out (i.e. keep) individuals who only used 1 boat in a year, and boats with
# only one permit holder fishing in a year.
cfecnew = cfec
#cfec.1boat = group_by(cfecnew, p_holder, year) %>%
#  summarize(nBoat = length(unique(cadfg))) %>%
#  filter(nBoat==1) %>%
#  left_join(cfecnew)
#cfecnew = cfec.1boat

dim(cfecnew)[1]
ndat <- list()
ndat$initial <- nrow(cfecnew)
```

## Combine some strategies

```{r}
codes <- readr::read_csv("data/cfec-codes.csv")
# convert 6-dig p_fshy codes to 5 digits
# removing blank space - now they're all length(5)
idx = which(nchar(codes$p_fshy)==6)
codes$p_fshy[idx] = paste(substr(codes$p_fshy[idx], 1, 1),
  substr(codes$p_fshy[idx], 3, 6), sep="")

# remove white space in 2nd position
idx = which(substr(codes$p_fshy, 2, 2) == " ")
codes$p_fshy[idx] = paste(substr(codes$p_fshy[idx], 1, 1),
  substr(codes$p_fshy[idx], 3, length(codes$p_fshy[idx])), sep="")

codes$description <- gsub(", VESSEL OVER 80", " VESSEL OVER 80", codes$description)
codes$description <- gsub(";97m", "", codes$description)
codes$description <- gsub(", VESSEL PERMIT OVER 80", " VESSEL PERMIT OVER 80", codes$description)
codes$description <- gsub(", VESSEL PERMIT TO 80", " VESSEL PERMIT TO 80", codes$description)
reg <- "([A-Za-z0-9/&(). -]+),([A-Za-z0-9()%/&.' -]+),([A-Za-z0-9/&.' ,()-]+)"
codes$pfshy_species <- tolower(gsub(reg, "\\1", codes$description))
codes$pfshy_gear <- tolower(gsub(reg, "\\2", codes$description))
codes$pfshy_area <- tolower(gsub(reg, "\\3", codes$description))
codes$pfshy_species <- sub(" +$", "", codes$pfshy_species)
readr::write_csv(codes, path = "portfolio/data-generated/cfec-codes.csv")
# as.data.frame(codes)
```

## Identify most popular strategies by person-year combination

Any person in a year will need to have 10% or more of their revenue to come
from each strategy (otherwise doesn't count as a strategy).

```{r}
cfecnew$p_spec = substr(cfecnew$p_fshy,1,1)
cfecnew$p_spec[cfecnew$p_spec == "B"] = "Halibut"
cfecnew$p_spec[cfecnew$p_spec == "C"] = "Sablefish"
cfecnew$p_spec[cfecnew$p_spec %in% c("D","E","K","T")] = "Crab"
cfecnew$p_spec[cfecnew$p_spec %in% c("F")] = "Freshwater"
cfecnew$p_spec[cfecnew$p_spec %in% c("G","H","L")] = "Herring"
cfecnew$p_spec[cfecnew$p_spec %in% c("I")] = "Lingcod"
cfecnew$p_spec[cfecnew$p_spec %in% c("A","J","N","R","W")] = "Shellfish"
cfecnew$p_spec[cfecnew$p_spec %in% c("O","P","Q","U","Z")] = "Invert"
cfecnew$p_spec[cfecnew$p_spec %in% c("M")] = "Misc"
cfecnew$p_spec[cfecnew$p_spec %in% c("S")] = "Salmon"
cfecnew$p_spec[cfecnew$p_spec %in% c("Y")] = "Rockfish"
cfecnew <- filter(cfecnew, p_spec != "Freshwater") # only 1400 rows
rev(sort(table(cfecnew$p_spec)))

permit_rev = group_by(cfecnew, p_holder, year, p_fshy) %>%
  summarize(permit.rev = sum(g_earn)) %>%
  as_data_frame() %>%
  group_by(p_holder, year) %>%
  mutate(tot.rev = sum(permit.rev))
permit_rev$permit.rev = permit_rev$permit.rev/permit_rev$tot.rev # fraction of total
ndat$rev_b4_10perc <- sum(permit_rev$permit.rev)
permit_rev = permit_rev[permit_rev$permit.rev < 0.1,]
ndat$rev_after_10perc <- sum(permit_rev$permit.rev)

# Delete these minimal earnings
nrow(cfecnew)
ndat$before_10_perc <- nrow(cfecnew)
cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% 
  paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy)), ]
nrow(cfecnew)
ndat$nrow_over_10_perc <- nrow(cfecnew)

topstrategies = group_by(cfecnew, p_holder, year) %>%
  summarize(strategy = paste(sort(unique(p_spec)), collapse = ", ")) %>%
  as_data_frame()
rev(sort(table(topstrategies$strategy)))[1:50]
cfecnew = left_join(cfecnew, topstrategies) # adds strategy column
write.csv(rev(sort(table(topstrategies$strategy))), file = "portfolio/data-generated/strategies.csv")

topstrategies.bypermit = group_by(cfecnew, p_holder, year) %>%
  summarize(strategy.permit = paste(sort(unique(p_fshy)), collapse = " ")) %>%
  as_data_frame()
cfecnew = left_join(cfecnew, topstrategies.bypermit) # adds strategy.permit column

# Summarize key species landings by strategy.permit -- this is for interpretation later
key.spec = c("SOCK","COHO","CHUM","PINK","CHNK","PCOD","PHLB","SABL","TCRB","KCRB","DCRB","PLCK","SAC","POUND")
species_by_strategy = group_by(cfecnew, strategy.permit, spec) %>%
  summarize(earn.bySpec = sum(g_earn)) %>% 
  data.frame %>% 
  group_by(strategy.permit) %>% 
  mutate(totEarn = sum(earn.bySpec)) %>% 
  filter(spec %in% key.spec)
species_by_strategy$earn.bySpec = species_by_strategy$earn.bySpec/species_by_strategy$totEarn
species_by_strategy = filter(species_by_strategy, earn.bySpec > 0.1) %>% select(-totEarn)
# convert to wide format
library(reshape2)
melted_sbs <- melt(species_by_strategy, id.vars = c("strategy.permit", "spec"))
species_by_strategy <- dcast(melted_sbs, strategy.permit ~ spec + variable)
write.csv(species_by_strategy, file = "portfolio/data-generated/species_by_strategy.csv")

```

Calculate gear based strategies:

```{r}
codes <- mutate(codes, p_gear = substr(codes$p_fshy, 2, 3))
gears <- select(codes, p_gear, pfshy_gear) %>% distinct()
cfecnew$p_gear = substr(cfecnew$p_fshy,2,3)
gear_freq <- sort(table(cfecnew$p_gear)) %>% as_data_frame()  %>% 
  rename(p_gear = Var1)
gears <- left_join(gear_freq, gears)
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
gears <- mutate(gears, pfshy_gear = trim(pfshy_gear))
gears <- group_by(gears, p_gear) %>% 
  summarize(n = n[1], pfshy_gear = paste(pfshy_gear, collapse = "-")) %>% 
    as_data_frame() %>% 
      arrange(-n)
gears %>% readr::write_csv(path = "portfolio/data-generated/gear-strategies.csv")

# hand edited with groups:
gears <- readr::read_csv("data/gear-strategies-edited.csv")
gears <- mutate(gears, p_gear = 
  ifelse(nchar(p_gear) == 1, paste0("0", p_gear), p_gear))

# adds gear group column:
cfecnew$gear_group <- NULL
cfecnew <- left_join(cfecnew, select(gears, p_gear, gear_group))

cfecnew$gear_group[cfecnew$p_spec == "Salmon" & 
  cfecnew$gear_group == "drift gillnet"] <- "drift gillnet salmon"
cfecnew$gear_group[cfecnew$p_spec == "Salmon" & 
  cfecnew$gear_group == "purse seine"] <- "purse seine salmon"
cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
  cfecnew$gear_group == "pot gear"] <- "pot gear sablefish/halibut"
cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
  cfecnew$gear_group == "longline under 60"] <- "longline under 60 sablefish/halibut"
cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
  cfecnew$gear_group == "longline over 60"] <- "longline over 60 sablefish/halibut"
cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
  cfecnew$gear_group == "longline over 90"] <- "longline over 90 sablefish/halibut"

topstrategies_gear <- group_by(cfecnew, p_holder, year) %>%
  summarize(strategy_gear = paste(sort(unique(gear_group)), collapse = ", ")) %>%
  as_data_frame()
rev(sort(table(topstrategies_gear$strategy_gear)))[1:50]

cfecnew <- left_join(cfecnew, topstrategies_gear) # adds gear strategy column

# xx <- filter(topstrategies_gear, strategy_gear == "")
# filter(cfecnew, p_holder == 138, year == 1990) %>% as.data.frame()
```

Aggregate annual metrics by person,

```{r}
cfecnew$yday <- paste(cfecnew$year, cfecnew$day, sep = "-") %>%
  lubridate::ymd() %>%
  lubridate::yday() - 1

if ("VLength" %in% names(cfecnew)) {
  cfecnew <- rename(cfecnew, vlength = VLength)
}

# calculate 
permit_days = group_by(cfecnew, p_holder, year, p_fshy) %>%
  summarize(permitDays = diff(range(yday)+1)) %>%
  group_by(p_holder, year) %>% 
  summarize(tot.permitDays = sum(permitDays))
cfecnew = left_join(cfecnew, permit_days)

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg[1],
    days = diff(range(yday) + 1),
    days_n = length(unique(yday)),
    days_permit = tot.permitDays[1], 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    weight = sum(g_pounds),
    length = mean(vlength, na.rm=TRUE),
    strategy = strategy[1],
    strategy_gear = strategy_gear[1],
    strategy_permit = strategy.permit[1],
    chnk = sum(g_earn[spec=="CHNK"]),
    coho = sum(g_earn[spec=="COHO"]),
    pink = sum(g_earn[spec=="PINK"]),
    sock = sum(g_earn[spec=="SOCK"]),
    chum = sum(g_earn[spec=="CHUM"]),
    region = region[1],
    nregion = length(unique(region)))

cfecAnnual = left_join(cfecAnnual, permit_days)
# also calculate weighted simpson diversity of species
cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, specn) %>%
  summarize(totRev.spec = sum(g_earn)) %>%
  group_by(p_holder, year) %>%
  summarize(specDiv = simp.div(totRev.spec)))

# also calculate weighted simpson diversity of gears
cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, gearn) %>%
  summarize(totRev.gear = sum(g_earn)) %>%
  group_by(p_holder, year) %>%
  summarize(gearDiv = simp.div(totRev.gear)))

# Also merge in previous revenue. Idx stores index of people last year
idx = match(paste(cfecAnnual$p_holder,cfecAnnual$year-1), paste(cfecAnnual$p_holder,cfecAnnual$year))

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
cfecAnnual$strategy_gear.prev = cfecAnnual$strategy_gear[idx]
cfecAnnual$strategy_permit.prev = cfecAnnual$strategy_permit[idx]
cfecAnnual$weight.prev = cfecAnnual$weight[idx]
cfecAnnual$days.prev = cfecAnnual$days[idx]
cfecAnnual$days_permit.prev = cfecAnnual$days_permit[idx]
cfecAnnual$chnk.prev = cfecAnnual$chnk[idx]
cfecAnnual$coho.prev = cfecAnnual$coho[idx]
cfecAnnual$chum.prev = cfecAnnual$chum[idx]
cfecAnnual$pink.prev = cfecAnnual$pink[idx]
cfecAnnual$sock.prev = cfecAnnual$sock[idx]

cfecAnnual$region.prev = cfecAnnual$region[idx]
cfecAnnual$nregion.prev = cfecAnnual$nregion[idx]
```

Save the data for modeling:

```{r}
# make sure the data are tidy:
cfecAnnual <- as_data_frame(cfecAnnual)
ndat$annual <- nrow(cfecAnnual)

saveRDS(cfecAnnual, file="portfolio/data-generated/cfec-annual-for-modeling.rds")

cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==FALSE, ]
saveRDS(cfecAnnual.diff, file="portfolio/data-generated/cfec-diff-for-modeling.rds")
ndat$annual_diff <- nrow(cfecAnnual.diff)
saveRDS(ndat, file = "portfolio/data-generated/ndat.rds")
```