### 2016-09-14 portfolio/text/supp.Rnw
Develop methods
Also moved some of the methods to the main paper

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{url}
\usepackage{mathtools}
\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\modulolinenumbers[2]

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
%\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\textheight 22.0cm

\usepackage[round,sectionbib]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

\setlength\parskip{0.10in}
\setlength\parindent{0in}

% Fix line numbering and align environment
% http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\title{Supporting Information}

\author{
}
\date{}

\begin{document}

\linenumbers

\setlength{\parindent}{0.9cm}
\setlength\parskip{0in}

\begin{spacing}{1.3}

<<set-knitr-options, cache=FALSE, echo=FALSE>>=
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
@

<<get-values, cache=FALSE, echo=FALSE>>=
load("../data-generated/output-values.rda")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
@

\maketitle

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\section{Filtering data and grouping permits into strategies}

We filtered the data in a number of ways before fitting our model.
We began with \Sexpr{round(ndat$initial/1e6,2)} million rows of data.
We removed revenue from permit-year combinations
that made up less than 10\% of annual revenue
for a given permit holder.
We removed permit holders whose median revenue
was less than \$10,000 USD (adjusted to 2009 values)
to focus on individuals actively engaged
in commercial fishing.
This reduced the total revenue from \Sexpr{round(ndat$rev_b4_large_only/1e9,0)}
billion 2009 USD
to \Sexpr{round(ndat$rev_large_only/1e9,0)} billion 2009 USD.

To limit the number of permits
and permit strategies
and focus on the main fisheries,
we removed all permits with less than 50 unique permit holders
across all years.
We also removed permits for freshwater species and
for experimental fisheries.
A list of all CFEC permit codes and descriptions is available at:
\url{https://www.cfec.state.ak.us/misc/FshyDesC.htm}

We then proceeded to combine similar permits in the following steps.
We combined a number of permits that were the same except for vessel size.
We combined permits for
otter troll,
longline finfish,
pot gear.
king crab,
and
shrimp.
We grouped a number of permits that were targeting a single species
and differed only in region fished.
For example, someone fishing herring roe in the Kodiak (G34K)
and someone fishing herring roe in Alaska Peninsula (G34M)
were both considered to have a herring roe permit strategy.
We combined permits across each of
dungeness crab,
herring roe purse seining,
herring roe gillnet,
herring hand picked,
herring pound fishing,
sea cucumber,
sea urchin,
tanner crabs

Combining salmon permits was more difficult because
the species composition varies widely based on geography.
We grouped salmon permits within gear type-region
combinations that shared similar species composition.
For the drift gillnet (S03) permits,
we combined permits from
Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
represents greater than 90\% of species revenue.
For the purse seine permits (S01),
there was a wide variety in species revenue by
area ($\sim$ 85\% pink in Prince William Sound to
85\% sockeye on the Alaskan Peninsula).
Of the five permits, only two had similar species compositions,
so we grouped Kodiak and Cook Inlet.
For the set gillnet salmon permits (S04),
there was some variety but also clear sockeye specialists.
We grouped permits that had greater than 80\% sockeye revenue.
This combined permits across
Prince William Sound,
Cook Inlet,
Kodiak,
Alaska Peninsula,
and Bristol Bay.
We grouped permits from Norton Sound
and Kuskokwim because of similar species compositions.
The remaining salmon permits were relatively unique
and were left ungrouped.

After this process of combining permits,
we were left with \Sexpr{n_grouped_permits}
unique fishing strategies (Table~\ref{tab:strategies}),
which we aggregated
within each person-year combination.
For example, if someone fished halibut
and sablefish in the same year
their strategy for that year would be ``halibut-sablefish''.
We removed people who switched strategies from one year to the next
since we were working with changes in revenue from year to year
and so needed consistent strategies across any pair of years
for a permit holder.
We removed strategies with less than 100 unique permit holders
across all years to focus on the major strategies
leaving us with a total of \Sexpr{nstr} strategies.
Finally, we removed all person-year combinations without
revenue recorded in the following year
in order to calculate percent change in revenue
from year to year.
This reduced the rows (person-year combinations)
in the final dataset from \Sexpr{ndat$annual} to
\Sexpr{ndat$annual_diff}.

TODO length(unique(pholders))
% Rows of data: \Sexpr{ndat$final}
% Number of permit holders: \Sexpr{npholders}
% Range of years: \Sexpr{range_yrs}
% Number of strategy-years: \Sexpr{nstr_yrs}
% \Sexpr{nstrat$enough_pholders}
% \Sexpr{round(ndat$rev_enough_pholders/1e9,2)}
% lost \% rev no strat:
% \Sexpr{round(ndat$rev_after_no_na_strat/ndat$rev_b4_no_na_strat,2)*100}
% lost rows no strat:
% \Sexpr{round(ndat$after_no_na_strat/ndat$b4_no_na_strat,2)*100}

\section{Hierarchical model}

We checked our models with a two-part modelling approach
using the R package lme4.
To do this, we fit the initial model of the mean component $\log(R_{i,t})$
and then fit the second model to the log absolute value of the
residuals, $\log | \widehat{\log(R_{i,t})} - \log(R_{i,t}) |$.
Most importantly, (1) this model does not carry uncertainty
from the first mean-component model to the variance-component model, but
(2) it also leaves the residuals in the mean-component model strongly
heteroskedastic and therefore improperly weights individual data points, and
(3) does not allow for proper modelling of the strategy-level predictors,
$\eta_1$ and $\eta_2$. Nonetheless, it provides a check that our results
from the Stan model are approximately correct and allowed us
to quickly iterate and build intuition about the models.
A version of the Stan model without the variance-component
matched the estimates of the same model in lme4 almost exactly.

% \verbatiminput{../analysis/portfolio-offset.stan}

\bibliographystyle{apalike}
\bibliography{refs}

\end{spacing}

\clearpage

\begin{table}[h!]
  \centering
  \caption{Labels and descriptions of fishing strategies. In many cases,
a particular strategy represents a group of similar permits across areas
or vessel sizes.}
  \label{tab:strategies}
  \begin{tabular}{ll}
  \toprule
\input{../figs/strategy-table.tex}
  \end{tabular}
\end{table}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=5in]{../figs/stan-main-effects.pdf}
\caption{Main effects from the hierarchical model
predicting revenue ($\beta$) and variability ($\gamma$) in revenue.
The $\eta$ coefficients represents strategy-level predictors
relating mean species diversity and mean days fished
within a strategy
(both centered by their means and divided by
two times their standard deivations)
to the magnitude of residual error.}
\label{fig:fe}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.06\textwidth]{../figs/stan-str-posteriors-dot.pdf}
\caption{Estimates of within-strategy effects from the hierarchical model.
Panels refer to parameters in Eqns \ref{eq:mean} and \ref{eq:cv}.
Shown are medians (dots) and 50\% and 95\% credible intervals
(thick and thin line segments).
Strategies are ordered from high to low mean
species diversity from top to bottom.}
\label{fig:re}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
\caption{The effect of adding a permit to ones strategy
on estimated revenue variability and median revenue.
This is an extended version of Fig.~2 that includes all major
strategy combinations.}
\label{fig:spoke-all}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid.png}
\caption{The log-absolute value residuals from the mean-component model (Eq.~\ref{eq:mean}).
It is these residuals that are modeled in the variance-component model
(Eq.~\ref{eq:cv}).
Breakpoint linear regression fits are overlaid in red for visual interpretation.}
\label{fig:resid}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid-down.png}
\caption{The log-absolute value of downside (negative) residuals from the main
component model.
Figure looks qualitatively the same as Fig.~\ref{fig:resid} and so justifies
using variability (positive and negative residuals) as a measure of risk.}
\label{fig:resid-downside}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{../figs/strategy-year-effects.pdf}
\caption{Strategy-year intercept estimates ($\beta_{0,j,t}$ in
Eq.~\ref{eq:mean}) from the hierarchical model.
Shown are 40 samples from the posterior distribution.}
\label{fig:str-yrs}
\end{center}
\end{figure}

% \begin{figure}[htbp]
% \begin{center}
% \includegraphics[width=0.9\textwidth]{../figs/lmer-vs-stan2.pdf}
% \caption{Comparison of }
% \label{fig:lmer-stan}
% \end{center}
% \end{figure}

\end{document}