### 2016-10-05 salmon/paper/salmon_paper_appendix.Rmd

---
title: ""
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
library(ggtern)
library(reshape2)
library(mgcv)
library(fpc)
library(gridExtra)
library(glmmTMB)
library(zoo)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
revenueThreshold = 10000

dat.annual = readRDS(file="../../portfolio/data-generated/cfec-annual-for-modeling.rds")

# single permit holders, salmon permits
dat.annual  = filter(dat.annual ,npermit==1 & substr(strategy_permit,1,1)=="S") %>%
  group_by(strategy_permit) %>% 
  mutate(nPeople = length(unique(p_holder))) %>% 
  filter(nPeople > 200) %>% 
  dplyr::select(-nPeople)
# drop troll
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S05B","S15B")==FALSE,]
# drop S04X, Y, Z because low participation (< 20) in many years
# S04W deleted because low median earnings
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S04W", "S04X","S04Y","S04Z")==FALSE,]

agg.annual = group_by(dat.annual, strategy_permit, year) %>% 
  summarize(people = length(unique(p_holder)),
    specDiv = mean(specDiv),
    meanrev = median(revenue),
    low = quantile(revenue,0.25),
    upp = quantile(revenue,0.75))

dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]
dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01L","S01M",
  "S03A","S03E","S03H","S03M","S03T","S04D","S04H","S04K","S04M","S04T"),]

agg.diff = group_by(dat.diff, strategy_permit) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
    cv = sd(log(revenue/revenue.prev)))

# Also load in raw CFEC data for plots
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
cfec$year = as.numeric(substr(cfec$landdate, 1, 4))

# only use people who are using just that permit - diversification within permit
cfec.raw = cfec
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]

# Restrict analysis to people who keep same strategy - complicates days, etc
dat.salmon = dat.salmon[which(dat.salmon$strategy_permit==dat.salmon$strategy_permit.prev),]

dat.salmon$days.change = log((dat.salmon$days_permit+1) / (dat.salmon$days_permit.prev+1))
dat.salmon$year_permit = as.factor(paste(dat.salmon$year, dat.salmon$strategy_permit))

cfec = cfec[which(paste(cfec$p_holder,cfec$year) %in% paste(dat.salmon$p_holder,dat.salmon$year)),]

# Make come plots for the appendix of (1) median fishing dates over time, (2) length of fishing season over time (calculated as width of 95% intervals)
cfec.raw$month = as.numeric(substr(cfec.raw$day, 1,2))
cfec.raw$day = as.numeric(substr(cfec.raw$day, 4,5))
cfec.raw$jday = mdy.date(cfec.raw$month, cfec.raw$day, cfec.raw$year) - mdy.date(1, 1, cfec.raw$year)
# summarize the temporal distribution of the catches
seasons = group_by(cfec.raw, p_fshy, year) %>% 
  summarize(q = mean(sample(jday, size = 10000, replace=T, prob=g_earn)),
    lower = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.025),
    upper = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.975),
    length = upper - lower)
seasons = rename(seasons, permit = p_fshy)
appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day")
appendixPlot0_p2 = ggplot(seasons, aes(year, length, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Season length")
```

```{r, echo = FALSE, eval=FALSE}
dat = dat.salmon
# create other category, lumping coho/chinook
dat$salm = dat$coho+dat$chnk
dat$salm.prev = dat$coho.prev+dat$chnk.prev

dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev

dat$salm = dat$salm/dat$totsalm
dat$sock = dat$sock/dat$totsalm
dat$pink = dat$pink/dat$totsalm
dat$chum = dat$chum/dat$totsalm
dat$chnk = dat$chnk/dat$totsalm
dat$coho = dat$coho/dat$totsalm
dat$salm.prev = dat$salm.prev/dat$totsalm.prev
dat$sock.prev = dat$sock.prev/dat$totsalm.prev
dat$pink.prev = dat$pink.prev/dat$totsalm.prev
dat$chum.prev = dat$chum.prev/dat$totsalm.prev
dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
dat$coho.prev = dat$coho.prev/dat$totsalm.prev

dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
dat$logdiff = log(dat$revenue/dat$revenue.prev)

dat_old = dat
dat$sumP2 = abs(dat$salm - dat$salm.prev) + 
  abs(dat$pink-dat$pink.prev) + 
  abs(dat$chum-dat$chum.prev) + 
  abs(dat$sock-dat$sock.prev)
dat = dat[which(dat$sumP2 < 0.3),]
```

```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere", cache=TRUE}

df = dat[which(dat$strategy_permit=="S01E"),]
pamk_s01e <- pamk(df[,c("chum","pink","sock","salm")])
df$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = group)) + geom_line() + ylab("People") + ggtitle("S01E")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = group)) + geom_line() + ylab("Mean revenue") + ggtitle("S01E")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = group)) + geom_line() + ylab("CV revenue") + ggtitle("S01E")

g = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T))

gridExtra::grid.arrange(p3,p0,p1,p2,ncol=2)
```

\break

```{r makePlots, echo = FALSE}
appendixPlot1 = ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Participation") + xlab("Year")

appendixPlot2_p1 = ggplot(agg.diff, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2)
appendixPlot2_p2 = ggplot(agg.diff, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2)
appendixPlot2_p3 = ggplot(agg.diff, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2)
```

\break

```{r, fig.cap="Species composition for 1988 and 2014 (each dot is a person) in the S03E fishery", fig.pos="placeHere", echo=FALSE, fig.height = 2, eval=FALSE}
### One way to visualize the change is with ternary plots over the 3 axes (chum / sockeye / pink)

daty = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","PINK","SOCK","COHO","CHNK"),], spec, p_holder, year) %>%
  summarize(p = sum(g_earn)) %>%
  group_by(p_holder, year) %>% mutate(p = p/sum(p))
daty <- melt(daty, id.vars = c("spec", "p_holder", "year"))
daty <- dcast(daty, p_holder + year ~ spec + variable)[,-1]
daty[is.na(daty)]=0
ggtern(data=daty[which(daty$year%in%c(1988,2014)),c("year", "CHUM_p","PINK_p","SOCK_p")], aes(x=CHUM_p,y=PINK_p,SOCK_p,color=as.factor(year))) +
  geom_point(alpha=0.3,size=1) +
  labs(x="CHUM",y="PINK",z="SOCK",title="Proportion of annual revenue (S03E)")
```

\break

```{r, echo=FALSE, eval=FALSE}

mu.mod = glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)

cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), dat.salmon)
# to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
# benefit (e.g. is the effect negative or positive), and (2) the magnitude (how steep / shallow it is). With the
# quadratic model, 
# 'squared term' indicates concave up/down
# derivative is magnitude, 2a * x + b

df = ranef(mu.mod)$cond$year_permit
colnames(df) = c("c","b","a")
df$year = as.numeric(substr(rownames(df), 1, 4))
df$strategy = substr(rownames(df), 6, 9)
df$deriv = 2*df$a + df$b
# also calculate change of going from 1 to 2 spp
df$pred1 = df$a*(1^2) + df$b*1 + df$c + fixef(mu.mod)$cond["specDiv.change"] + fixef(mu.mod)$cond["specDiv.change:specdiv.prev"]*1
# add fixed effects into predictions

# We can demonstrate the quadratic curves by year for a single strategy / set of years
df.s01e = df[df$strategy=="S01E",]
demo = expand.grid("speciesDiversity" = seq(1,4,length.out=100), "year" = 1986:2014)
demo = left_join(demo, df.s01e)
demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
```

```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide', eval=FALSE}
pdf("../../salmon/Salmon_residuals.pdf")
# Residuals for mean model
dat.salmon$mu.resid = residuals(mu.mod)
dat.salmon$mu.fitted = fitted(mu.mod)

ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specDiv.change, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")

# Residuals for CV model
dat.salmon$cv.resid = residuals(cv.mod)
dat.salmon$cv.fitted = fitted(cv.mod)

ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
dev.off()
```
```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Variation in effects of diversification by year for the S01E (PWS purse seine) permit holders. ", eval=FALSE}
###We can make quadratic plots of these effects by year and strategy. These plots represent the effect of increasing diversity (as relative change) on mean revenue. For example, in 1986 if you were fishing 3-4 species there would be little effect of adding more spp; but if you were fishing 1 species, there would be a slight negative effect of adding more. In 1991, there would be a huge benefit of increasing diversity, and that effect would taper off as more were added.

ggplot(demo, aes(speciesDiversity, effect, group = year)) + geom_line() + facet_wrap(~ year) + xlab("Species Diversity") + ylab("Effect of increasing diversity (relative change)")
```

\break

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3, eval=FALSE}
###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue. For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010. 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse. Across permits, perhaps slight declines in some of these in recent years?

ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3, eval=FALSE}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit.", eval=FALSE}
df.cv = ranef(cv.mod)$cond$year_permit
colnames(df.cv) = c("c.cv","b.cv","a.cv")
df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
df.cv$strategy = substr(rownames(df.cv), 6, 9)
# also calculate change of going from 1 to 2 spp
df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1

ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")

```

\break

```{r, echo=FALSE, eval = FALSE}
# These are the run  timing plots per Milo's questions

# for 2008, create cumulative density plots of salmon landings by calday
pinkchum = cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec%in%c("PINK","CHUM","SOCK"),]
#pinkchum = cfec.raw[cfec.raw$p_fshy=="S 03T" & cfec.raw$spec%in%c("PINK","SOCK","CHNK"),]

# aggregate by calendar day
pinkchum$month = as.numeric(substr(pinkchum$day, 1, 2))
pinkchum$day = as.numeric(substr(pinkchum$day, 4, 5))
pinkchum$jday = date::mdy.date(pinkchum$month, pinkchum$day, pinkchum$year) - date::mdy.date(1,1,pinkchum$year)

group_by(pinkchum, jday, year) %>% 
  summarize(pink = sum(g_earn[spec=="PINK"]), 
    chum = sum(g_earn[spec=="CHUM"]),
    sock = sum(g_earn[spec=="SOCK"])) %>% 
  group_by(year) %>% 
  mutate(tot = sum(pink) + sum(chum) + sum(sock), pink = pink/tot, chum = chum/tot, sock = sock/tot) %>% 
ggplot(aes(jday, pink)) + geom_line(color="pink") + geom_line(aes(jday, chum), color="darkgreen") + geom_line(aes(jday, sock), color="lightblue") + facet_wrap(~ year, scale="free")

# make plot of all earnings or weight by jday
salm = cfec.raw[cfec.raw$p_fshy=="S 01E",]
salm$month = as.numeric(substr(salm$day, 1, 2))
salm$day = as.numeric(substr(salm$day, 4, 5))
salm$jday = date::mdy.date(salm$month, salm$day, salm$year) - date::mdy.date(1,1,salm$year)

g = group_by(salm, jday, year) %>% 
  summarize(earn = sum(g_earn), 
    lbs = sum(g_pounds), 
    pink = sum(g_earn[spec=="PINK"])/earn,
    sock = sum(g_earn[spec=="SOCK"])/earn,
    chum = sum(g_earn[spec=="CHUM"])/earn,
    coho = sum(g_earn[spec=="COHO"])/earn,
    chnk = sum(g_earn[spec=="CHNK"])/earn) %>% 
ggplot(g, aes(jday, (earn))) + geom_line(colour = rgb(g$sock+g$coho+g$chnk, g$chum, g$pink)) + facet_wrap(~ year, scale="free_y")

```

```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere", eval=FALSE}

df = dat[which(dat$strategy_permit=="S01A"),]
pamk_s01a <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01a$pamobject$clustering

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle("S01A")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle("S01A")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle("S01A")

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

\pagebreak

# Appendix

## Overall trends in participation across permits

```{r, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
appendixPlot1
```

\break

```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
  summarize(gearn = sum(g_earn)) %>%
  group_by(year, p_fshy) %>%
  summarize(med = median(gearn), low = quantile(gearn, 0.25),
    upp = quantile(gearn,0.75)) %>%
  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy,scale="free") + 
  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue") + xlab("Year")
```

```{r, echo=FALSE, fig.height = 3.1, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
group_by(cfec, p_fshy, year, p_holder) %>% 
  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(meanW = mean(meanW, na.rm=T)) %>% 
  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons / person") + xlab("Year") + ggtitle("Trends in landed weight")
```

\break

```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
```

\break

## Species composition

```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
# plot proportion of species landed by salmon permit
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() + facet_wrap(~p_fshy) + 
  ylab("Species % (total) ") + xlab("")
```

```{r, fig.cap="Effective species diversity across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
group_by(dat.salmon, year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + facet_wrap(~strategy_permit, scale="free") + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Effective diversity (across people)") + xlab("")
```

```{r, fig.cap="Proportion of total returns to Prince William Sound (hatchery + wild, catch + escapement). Years before 1980 and after 2011 aren't included because not all species have data.", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
returns = readxl::read_excel("salmon_returns_PWS.xlsx")
returns = returns[,c("Year","Sockeye_total","Pink_total","Chum_total","Coho","Chinook")]

returns = as.data.frame(returns[which(returns$Year%in%c(1980:2011)),])
for(i in 1:ncol(returns)) returns[,i] = as.numeric(returns[,i])
for(i in 1:nrow(returns)) {
    # normalize
    returns[i,-1] = returns[i,-1] / sum(returns[i,-1])
}

melted = reshape2::melt(returns,id.vars="Year")
names(melted)[which(names(melted)=="variable")]="species"
ggplot(melted,aes(Year,value, group = species, color=species)) + geom_line() + ylab("Proportion of total returns (by #)")
```

\break

## Trends in revenue and prices

Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook. Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.  For most permit types, median revenue / person has generally increased, especially since early 2000s. For some permits, variability appears to have changed through time (quartile range getting smaller).

```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
# plot prices of species landed by salmon permit
cfec.raw$area = NA
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"

group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
  ylab("Revenue (millions) ") + xlab("Year")
```

\break

## Trends in variability

```{r, fig.cap="Trends in annual variability by fishery", fig.pos="placeHere", echo=FALSE,  fig.height=3.3}
summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, spec) %>% 
  summarize(sumg = log(sum(g_earn)), sumw = log(sum(g_pounds))) %>% 
  arrange(p_fshy, spec, year) %>% group_by(p_fshy, spec) %>% 
  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA)), 
    logdiffw = rollapply(c(NA, diff(sumw)), 3, sd, fill=c(NA, NULL, NA)))

ggplot(summary, aes(year, logdiffg, color=spec)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("3 year rolling CV (revenue)") + xlab("Year")
```

```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere",  echo=FALSE, fig.height=3.3}
g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Pink salmon")

g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Chum salmon")

gridExtra::grid.arrange(g1,g2, ncol=1)
```

```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere", echo=FALSE, eval=FALSE}
g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Pink salmon")

g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Chum salmon")

gridExtra::grid.arrange(g1,g2, ncol=1)
```

```{r, fig.cap="Trends in variability across people", fig.pos="placeHere", echo=FALSE, fig.height=3.5, eval=FALSE}
summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
  summarize(sumg = sum(g_earn), sumw = sum(g_pounds)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(cvg = sd(sumg,na.rm=T)/mean(sumg,na.rm=T),
    cvw = sd(sumw,na.rm=T)/mean(sumw,na.rm=T))
ggplot(summary, aes(year, cvg )) + geom_line(color ="green" ) + facet_wrap(~ p_fshy, scale="free") + geom_line(aes(year, cvw), color = "blue") + ylab("3 year rolling CV") + xlab("Year") + ggtitle("Variability across people: total revenue (green) and weight (blue)")
```

```{r, fig.cap="Change in CV for individual people", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
  summarize(sumg = sum(g_earn)) %>% 
  group_by(p_fshy, p_holder) %>% 
  mutate(nPre = length(which(year%in%seq(1985,1995))),
    nPost = length(which(year%in%seq(2005,2014)))) %>% 
  filter(nPre > 3 & nPost > 3) %>% select(-nPre, -nPost) %>% 
  summarize(cv.pre = mean(sumg[which(year <= 1995)],na.rm=T)/sd(sumg[which(year <= 1995)],na.rm=T),
    cv.post = mean(sumg[which(year >= 2005)],na.rm=T)/sd(sumg[which(year >= 2005)],na.rm=T))

ggplot(summary, aes(log(cv.post/cv.pre))) + geom_histogram(bins=100) + facet_wrap(~ p_fshy) + geom_vline(aes(xintercept = 0)) + ylab("Count") + xlab("% Change in CV for individual people (2005-2014 / 1985-1995)") + ggtitle("Change in variability for individual people")
#summary = left_join(expand.grid(year = unique(summary$year), p_holder = unique(summary$p_holder)), summary)

# add rolling CV for each person
#summary = group_by(summary, p_holder) %>% 
#  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA))) %>%
#  do(fit = lm(logdiffg ~ year, data = .))
```

\break

```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}

g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
p1 = group_by(g, spec, area) %>% 
  mutate(p = p/p[which(year==2002)]) %>% 
  group_by(year, area) %>%
  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change") + xlab("")
p1

```

```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}

g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
# do log-diff of prices -- get price year before
g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
  paste(g$year+1, g$spec, g$area))]

p2 = group_by(g, spec, area) %>% 
  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") + xlab("")
p2

g$logdiff = log(g$p/g$p_t1)
p3 = group_by(g, year, area) %>% 
summarize(sd = sd(logdiff,na.rm=TRUE)) %>% ggplot(aes(year, sd, color=area)) + geom_line() + xlab("Year") + ylab("SD of price changes across species") + ggtitle("Synchrony in price fluctuations across species")

```

```{r, fig.pos="placeHere", fig.cap = "Standard deviation (across species, by area) for fluctuations in prices. Fluctuations are calculated as the log difference price data between sequential years. The standard deviation of log differenced data is interpreted as the coefficient of variation (CV)."}
p3
```

\break

### Season length

There appears to be no significant trend in season length (defined as the difference between 95% CIs) or mean julian day for any fishery. In all cases, the dates are weighted by the earnings associated with landed catch.

```{r, fig.cap="Trends in mean julian day and season length by fishery", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
gridExtra::grid.arrange(appendixPlot0_p1, appendixPlot0_p2, ncol = 1)
```