### 2016-11-10 diversity/analysis/herring_case_study.Rmd
herring case study

---
title: "Herring: case study 1"
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
library(reshape2)
library(cluster)
library(feather)
```

## The Boom and Bust of Pacific Herring

Pacific herring (*Clupea pallasii*) have been one of the focal species associated with the Exxon Valdez Oil Spill (EVOS) (Carls et al. 2002, Thorne and Thomas 2008, )

have been one of the most important commercial and subsistence resources in the history of Alaska, with commercial fisheries established in 1878 (Woodby et al. 2005). Herring were initially harvested for oil in the 1880s at Killisnoo (U.S. Department of Interior 1893; Skud et al. 1960). For much of the next 50 years, the herring reduction fishery continued to expand, producing both bait and oil. Statewide, the herring reduction harvest was highest in 1937 and declined to the mid-1960s (Reid 1971). Prince William Sound experienced similar dynamics, with commercial harvest beginning as a bait fishery in 1913 (Rounsefell and Dahlgren 1935), and the reduction fishery peaking in 1937 (Funk 1994). Unlike more recent herring fisheries, these bait and reduction fisheries generally targeted schools of herring in the ocean. Over time, their size-selectivity became smaller, and they started targeting smaller fish for reduction (Funk 1994). In the 1950s, less than 10 boats were participating in the reduction fishery, and the continued decline in herring abundance, combined with cheaper altenratives from foreign markets (Peruvian anchovetta) led to the closure of the fishery in the late 1960s (Funk 1994).

## Shifting demand: the graying of the palette

In the early 1970s, new herring fisheries targeting a new herring product once again emerged in Prince William Sound. Declines of herring in Russia and Japan (Carlson 2005; Woodby et al. 2005) prompted increased demand for salted herring roe (and more recently, luxury products like spawned herring eggs on kelp). Unlike the reduction fisheries, herring roe fisheries targeted herring near shore immediately before they spawned (spring), using a variety of gears, including gillnets, beach seines, and purse seines.

Trends in revenue and landings in Prince William Sound fisheries largely mirror those of the rest of the state, with peaks in 1988 (Carlson 2005). Coincident with EVOS in 1989, harvest and prices  and declines through the 1990s.

In addition to being susceptible to changing demand, a related pressure on Alaskan sac roe fisheries has been external markets. Over the period 1994-2003 for example, roe from Alaska represented a small fraction of the total supplied to Japan, especially compared to roe from Canada and China (Carlson 2005).

## Changing Participation

Declines in herring abundnace, demand (and prices) and ex-vessel values in Prince William Sound have been well documented (Carlson 2005). Following peak

Participation

\break

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Fishery landings for herring in Prince William Sound. L12E and L21E permits are combined for spawn on kelp, H01E and H34E are combined for food and bait, and G01E and G34E are combined for sac roe fisheries."}

cfec <- read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

pws = dplyr::filter(cfec, p_fshy%in%c("L 12E", "L 21E", "G 01E", "G 34E", "H 01E", "H 34E"))
pws$p_fshy[pws$p_fshy%in%c("L 12E","L 21E")] = "Spawn on kelp" 
pws$p_fshy[pws$p_fshy%in%c("H 01E","H 34E")] = "Food and bait"
pws$p_fshy[pws$p_fshy%in%c("G 01E","G 34E")] = "Roe"

names(pws)[which(names(pws)=="p_fshy")]="Fishery"
g1 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
  summarize(w = sum(g_pounds)) %>% 
  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
  xlab("Year") + ylab("Pounds")

g2 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
  summarize(w = length(unique(p_holder))) %>% 
  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
  xlab("Year") + ylab("Active permits")

gridExtra::grid.arrange(g1, g2)
```

## Changing Strategies

```{r, eval=FALSE}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

this.permit = "G 01E"
# S01E S03T - effective permit diversity

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}
```

```{r, eval=FALSE}
g1 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>%
  group_by(p_holder, year) %>% 
  summarize(div = simp.div(g)) %>% 
  group_by(year) %>% 
  summarize(mean = mean(div), lower=quantile(div,0.025), upper=quantile(div,0.975)) %>% 
  ggplot(aes(year, mean)) + geom_line() + ggtitle(this.permit) + ylab("Mean diversity")

# Identify dominant strategies associated with this permit
g2 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>%
  group_by(p_holder, year) %>% 
  mutate(totRev = sum(g)) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  mutate(g = g / totRev) %>% 
  select(-totRev)

# use reshape 2 to cast this to wide format
aql <- melt(g2, id.vars = c("year", "p_fshy", "p_holder"))
aqw <- dcast(aql, year + p_holder ~ p_fshy)
# replace NAs w/0s
for(i in 1:ncol(aqw)) {
  aqw[which(is.na(aqw[,i])),i] = 0
}

indx = sample(seq(1,nrow(aqw)),size=500,replace=F)
sub = aqw[indx,-c(1,2)]

sub = aqw[,-c(1,2)]
coarse_df = aqw[,-c(1,2)]
coarse_df$S = apply(coarse_df[,which(substr(names(sub),1,1)=="S")],1,sum)
coarse_df$B = apply(coarse_df[,which(substr(names(sub),1,1)=="B")],1,sum)
coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))],1,sum)
coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
coarse_df$P = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("P"))],1,sum)
coarse_df = coarse_df[,c("S","B","GHL","KTD","M","P")]

clust = fpc::pamk(coarse_df[sample(seq(1,nrow(aqw)),size=500,replace=F),])# cluster based on subset to pick medoids
medoids = clust$pamobject$medoids
clust = pam(coarse_df, k = clust$nc) # cluster full dataset

names_medoids = ""
for(i in 1:nrow(medoids)) {
  names_medoids[i] = paste0(colnames(medoids)[which(medoids[i,] > 0.05)], collapse=":")
}

aqw$clusters = clust$clustering
totals = group_by(aqw, clusters, year) %>% 
  summarize(n = n())
totals$group = names_medoids[totals$clusters]

ggplot(totals, aes(year, n, group=group, color=group)) + geom_line() + geom_point()

ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1)

```

```{r,eval=FALSE}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

```

```{r eval=FALSE}
# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", eval=FALSE}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

herring = cfec[substr(cfec$p_fshy,1,1) %in% c("L","G","H"),]

herring_summary = group_by(herring, p_fshy, year) %>%
  summarize(g = sum(g_earn), p = mean(g_price,na.rm=T), lbs = sum(g_pounds), people = length(unique(p_holder))) %>%
  group_by(p_fshy) %>%
  mutate(n = length(unique(year)), maxpeople = max(people,na.rm=T)) %>%
  filter(n > 9) %>%
  filter(maxpeople > 25)

quitters = group_by(herring, p_holder, p_fshy) %>%
  summarize(lastYear = max(year))

group_by(quitters[quitters$lastYear<2014,],lastYear,p_fshy) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(lastYear,n)) + geom_line() + facet_wrap(~p_fshy, scale="free_y") + xlab("Year") + ylab("People leaving the fishery (last year fished)")

# look at what these people did after leaving
cfec_quitters = cfec[which(cfec$p_holder%in%unique(quitters$p_holder)),]
cfec_quitters = left_join(cfec_quitters, quitters[,c("p_holder","lastYear")])

fates = as.data.frame(group_by(cfec_quitters, p_holder, p_fshy) %>% 
  summarize(n = length(which(year > lastYear[1]))) %>% 
  group_by(p_fshy) %>% 
  summarize(n0 = length(which(n==0)), nOther = length(which(n>0))))

trackQuitters = function(permit) {
cfec2= cfec_quitters[which(cfec_quitters$p_holder %in% unique(quitters[quitters$p_fshy==permit,]$p_holder)),]
cfec2 = cfec2[which(cfec2$year > cfec2$lastYear),]
cfec2$SPEC = substr(cfec2$p_fshy,1,1)
totalpermit = group_by(cfec2,SPEC) %>% 
  summarize(g = sum(g_earn))
totalpermit$g = totalpermit$g/sum(totalpermit$g)
return(as.data.frame(arrange(totalpermit,-g)))

}

#trackQuitters("L 21C")
```

## Overview

First, we can pull out the herring permits by all areas. We'll restrict the analysis to (1) fisheries with 10 or more years of data, and (2) fisheries that at their max had at least 25 people landing catch in a year. We see that people generally decline to 0, except fisheries in SE (G01A, G34A), and spawn on kelp fisheries (L21). Despite catch being flat or maybe increasing, we see prices go down or flat for nearly all permits (and spawn on kelp L21 permits have significantly higher price / lb compared to roe fisheries).

In terms of the diversification of these permit holders, some questions to ask might be (1) if people stopped fishing herring, did they stop fishing entirely, or fish something else? (2) for the people that stayed, how many of them are diversified (other permits) and how much of their revenue comes from other permits?

For a lot of permits, we see decreases in the proportion of people fishing only herring. In terms of spatial diversity, for several of the fisheries that haven't crashed, we see increases in diversity: G01M/T, G34N, L21. In terms of diversifying to other types of herring permits,

```{r, eval=FALSE, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Herring permit holders by year."}
  ggplot(herring_summary, aes(year, people)) + geom_line() + facet_wrap(~ p_fshy, scale="free_y") + xlab("Year") + ylab("People")
```

\break

```{r, eval=FALSE, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in landed catch by permit.", fig.height=3.5}
  ggplot(herring_summary, aes(year, log(lbs))) + geom_line() + 
  facet_wrap(~ p_fshy) + ylab("Ln (tons)") + xlab("Year")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in prices by permit.", fig.height=3.5}
  ggplot(herring_summary, aes(year, p)) + geom_line() + 
  facet_wrap(~ p_fshy, scale="free_y") + ylab("Price (Dollars / lb.)") + xlab("Year")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Scatterplots of prices versus revenue by permit."}
  ggplot(herring_summary, aes(p, log(lbs/2000), colour = year)) + geom_point() + facet_wrap(~ p_fshy, scale="free") + ylab("Ln (tons)") + xlab("Price (dollars / lb.)")

```

\break

## How much of your revenue is driven by non-herring?

```{r,eval=FALSE, echo = FALSE}

herring_permits = unique(herring_summary$p_fshy)

# the complicated thing here is that people may hold multiple herring permits, 
# multiple permits, no permits. So for each permit type we need to calculate 
# each of these groups independently. Some people may be counted multiple times.

# next calculate strategies for diversification
calc_permit_summary = function(this_permit) {
people_year = group_by(cfec[cfec$p_fshy == this_permit,], p_holder, year) %>% 
  summarize(g = sum(g_earn)) %>% 
  filter(g > 0) %>% 
  select(-g)
# grab all records pertaining to these people-year combinations
herring = left_join(people_year, cfec)

# aggregate fish tickets to permits
herring = group_by(herring, p_holder, p_fshy, year) %>% 
  summarize(g_earn = sum(g_earn))

permit_summary = group_by(herring, p_holder, year) %>% 
  summarize(npermit = length(unique(p_fshy)),
    npermit_herring = length(which(p_fshy %in% herring_permits)),
    npermit_g34 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 34")),
    npermit_g01 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 01")),
    npermit_l12 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 12")),
    npermit_l21 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 21")),
    nmult = ceiling(npermit_g34/100) + ceiling(npermit_g01/100) + ceiling(npermit_l21/100) + ceiling(npermit_l12/100)) %>% 
    as.data.frame() %>%
    group_by(year) %>% 
  summarize(only_herring_single = length(which(npermit==1 & npermit_herring == 1)),
    only_herring_multiple = length(which(npermit>1 & npermit_herring >1 & npermit == npermit_herring)), herring_combo = length(which(npermit > 1 & npermit_herring >= 1 & npermit > npermit_herring)), g34 = length(which(npermit_g34 > 1)), g01 = length(which(npermit_g01 > 1)),
    l12 = length(which(npermit_l12 > 1)), l21 = length(which(npermit_l21 > 1)), 
    multiple_fisheries = length(which(nmult > 1)))

permit_summary$p_fshy = this_permit
return(permit_summary)
}

herring_summary = calc_permit_summary(herring_permits[1])
for(i in 2:length(herring_permits)) {
  herring_summary = rbind(herring_summary, calc_permit_summary(herring_permits[i]))
}

```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders specializing on herring.", fig.height=3.5}
ggplot(herring_summary, aes(year, (only_herring_single+only_herring_multiple)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people only fishing herring")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders diversifying spatially.", fig.height=3.5}
p1 = ggplot(herring_summary, aes(year, (g34)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p2 = ggplot(herring_summary, aes(year, (g01)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p3 = ggplot(herring_summary, aes(year, (l21)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p4 = ggplot(herring_summary, aes(year, (l12)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

gridExtra::grid.arrange(p1, p2, p3, p4)
```

\break

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of people diversifying w/diff herring permits.", fig.height=3.5}
ggplot(herring_summary, aes(year, (multiple_fisheries)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying w/diff herring permits")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Ratio of effective diversity to total number of permit holders (as measure of evenness). Values near 1 indicate everyone's revenue is similar.", fig.height=3.5}
# We can also calculate how herring catch is distributed across fisherment
simp.div = function(x) {1/sum((x/sum(x))^2)}

group_by(cfec[cfec$p_fshy%in%herring_permits,], year, p_holder, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>% 
  group_by(year, p_fshy) %>% 
  summarize(div = simp.div(g), n = length(unique(p_holder))) %>% 
  ggplot(aes(year, div/n)) + facet_wrap(~p_fshy, scale="free_y") + geom_line() + xlab("Year") + ylab("Diversity across people")
```

\break

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Fraction of revenue made up of herring (median, of per-person fractions).",eval=FALSE}

herring_permits = unique(herring_summary$p_fshy)

herring_people = group_by(cfec, p_fshy, year, p_holder) %>%
  summarize(g = sum(g_earn),
  herring_landing = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1)) %>% 
  data.frame() %>% 
  group_by(p_holder, year) %>% 
  mutate(herring_personyear = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1),
  g_herring = sum(g[herring_landing==1]), g_other = sum(g[herring_landing==0]),
  p_fshy = p_fshy[which(p_fshy %in% herring_permits)][1]) %>% 
  group_by(p_holder) %>% 
  mutate(ever_herring = sum(herring_personyear)) %>%   
  filter(ever_herring > 0) %>% 
  select(-g, -herring_landing, -herring_personyear, -ever_herring)

herring_people$p_herring = herring_people$g_herring / (herring_people$g_herring + herring_people$g_other)
group_by(herring_people[herring_people$p_fshy %in% herring_permits, ], p_fshy, year) %>% 
  summarize(g = median(p_herring)) %>% 
  ggplot(aes(year, g)) + geom_line() + facet_wrap(~p_fshy) + ylab("Median % herring (as fraction of total revenue)") + xlab("Year")
```

## References

Carls, M.G., G.D. Marty, and J.E. Hose. 2002. Synthesis of the toxicological impacts of the Exxon Valdez oil spill on Pacific herring (Clupea pallasi) in Prince William Sound, Alaska, U.S.A. Canadian Journal of Fisheries and Aquatic Sciences, 2002, 59(1): 153-172.

Carlson, S. 2005. Changes in roe herring markets: a review of available evidence. CFEC report number 05-5N. Commercial Fisheries Entry Commission, Juneau, AK.

Funk, F. 1994. Forecast of the Pacific herring biomass in Prince William Sound, Alaska, 1993. Regional Information Report, 5J94-04. ADFG, Commercial Fisheries and Managmenet Division.

Reid, G.M. 1971. Age composition, weight, length, and sex of herring, Clupea pallasii, used for 
reduction in Alaska, 1929-1966.

Rounsefell, G.A. and E.H. Dahlgren. 1935. Fluctuations in the supply of herring, Clupea pallasii, in Prince William Sound, Alaska. Bulletin of the U.S. Bureau of Fisheries 47:263-291.

Skud, B.E., H.M. Sakuda, and G.M Reid. 1960. Statistics of the Alaska herring fishery, 1878-1956. U.S. Fish and Wildlife Service, Statistical Digest 48, 21 p.

Thorne, R. E., and Thomas, G. L. 2008. Herring and the "Exxon Valdez" oil spill: an investigation into historical data conflicts. – ICES Journal of Marine Science, 65: 44–50.

U.S. Department of Interior 1893. Population and Resources of Alaska at the Eleventh Census: 1890. U.S. Department of Interior, Census Office.

Woodby, D., D. Carlile, S. Siddeek, F. Funk, J.H. Clark, and L. Hulbert. 2005. Commercial Fisheries in Alaska. ADFG special publication 05-09.  http://www.sf.adfg.state.ak.us/FedAidPDFs/sp05-09.pdf