### 2017-01-06 portfolio/text/anderson-etal-rev-pfx.tex
Jordan and Milo edits

\documentclass[12pt,english]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabu} % for spacing between rows in longtable
\usepackage{titlesec}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{commath} % for \abs{}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{4pt}\setlength{\parskip}{0pt}}

\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\modulolinenumbers[2]

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\textheight 22.0cm

% \usepackage[round,sectionbib]{natbib}
% \bibpunct{(}{)}{;}{a}{}{;}

% Fix line numbering and align environment
% http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\title{
  Benefits and risks of diversification\\ for individual fishers
  % Benefits of diversification and risks of 
  % changing strategies for individual fishers\\
% Benefits of diversification are unavailable to many fishermen\\
}
\author{
Sean C. Anderson$^1$,
Eric J. Ward$^2$,
Andrew O. Shelton$^2$,\\
Milo D. Adkison$^3$,
Anne H. Beaudreau$^3$,\\
Richard E. Brenner$^4$,
Alan C. Haynie$^5$,
Jennifer C. Shriver$^4$,
Jordan T. Watson$^6$,
\ldots
}

\newlabel{tab:strategies}{1}
\newlabel{fig:fe}{1}
\newlabel{fig:re}{2}
\newlabel{fig:spoke-all}{3}
\newlabel{fig:resid}{4}
\newlabel{fig:resid-downside}{5}
\newlabel{fig:str-yrs}{6}
\newlabel{fig:ifq}{7}
\newlabel{fig:mirror}{8}
\newlabel{fig:noeffort}{9}

\date{}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\begin{document}
% \begin{flushleft}

\maketitle

% TODO: standardize and complete addresses

\noindent
$^1$School of Aquatic and Fishery Sciences, Box 355020,
University of Washington, Seattle, WA 98195, USA

\noindent
$^2$Conservation Biology Division, Northwest Fisheries Science Center,
National Marine Fisheries Service, National Oceanographic and
Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA

\noindent
$^3$College of Fisheries and Ocean Sciences, University of Alaska Fairbanks,
17101 Point Lena Loop Rd, Juneau, AK, 99801, USA

\noindent
$^4$Alaska Department of Fish and Game, Commercial Fisheries Division, Cordova, USA

\noindent
$^5$Alaska Fisheries Science Center,
National Marine Fisheries Service, National Oceanographic and
Atmospheric Administration, Bldg 4,
7600 Sand Point Way NE, Seattle, WA, 98115, USA

\noindent
$^6$NOAA / National Marine Fisheries Service, Alaska Fisheries Science Center, Auke
Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA

\setlength{\parindent}{0.9cm}
\setlength\parskip{0in}

\thispagestyle{empty}

\clearpage

\linenumbers

\begin{spacing}{1.4}
\input{ms}

\end{spacing}
% \end{flushleft}
\end{document}