### 2017-01-13 portfolio/text/anderson-etal-rev-pfx.tex
Format paper

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Benefits and risks of diversification for individual fishers}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Sean C. Anderson}
\author[b]{Eric J. Ward}
\author[b]{Andrew O. Shelton}
\author[c]{Milo D. Adkison}
\author[c]{Anne H. Beaudreau}
\author[d]{Richard E. Brenner}
\author[e]{Alan C. Haynie}
\author[d]{Jennifer C. Shriver}
\author[f]{Jordan T. Watson}
\author[d]{Benjamin C. Williams}

\affil[a]{School of Aquatic and Fishery Sciences, Box 355020,
University of Washington, Seattle, WA 98195, USA}

\affil[b]{Conservation Biology Division, Northwest Fisheries Science Center,
National Marine Fisheries Service, National Oceanographic and
Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA}

\affil[c]{College of Fisheries and Ocean Sciences, University of Alaska Fairbanks,
17101 Point Lena Loop Rd, Juneau, AK, 99801, USA}

\affil[d]{Alaska Department of Fish and Game,
Division of Commercial Fisheries,
PO BOX 115526,
Juneau, AK, USA 99811-5526, USA}

\affil[e]{Alaska Fisheries Science Center,
National Marine Fisheries Service, National Oceanographic and
Atmospheric Administration, Bldg 4,
7600 Sand Point Way NE, Seattle, WA, 98115, USA}

\affil[f]{NOAA / National Marine Fisheries Service, Alaska Fisheries Science Center, Auke
Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA}

% Please give the surname of the lead author for the running footer
\leadauthor{Anderson}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{TODO}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
S.C.A., TODO designed the study; S.C.A., TODO analyzed the
data and wrote the paper with input from all authors.}
\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sandrsn@uw.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{
TODO $|$
TODO
}

\newcommand{\tabstrategies}{1}
\newcommand{\figfe}{1}
\newcommand{\figre}{2}
\newcommand{\figspokeall}{3}
\newcommand{\figresid}{4}
\newcommand{\figresiddownside}{5}
\newcommand{\figstryrs}{6}
\newcommand{\figifq}{7}
\newcommand{\figmirror}{8}
\newcommand{\fignoeffort}{9}

% 250:
\begin{abstract}
Individuals relying on the extraction of natural resources for their
livelihood---fishers, farmers, miners, and loggers, among others---face high
levels of income variability driven by a mix of environmental, biological,
management, and
economic dynamics. A key component to managing these industries is identifying
how regulatory actions and individual behaviour affect income variability and by
extension economic stability and the sustainable use of natural resources. For
commercial fisheries, communities and vessels fishing a greater diversity of
species have less revenue variability than those fishing a lower diversity.
However, it is unclear if these benefits extend to the actions of individual
fishers and how year-to-year changes in diversification affect revenue and
revenue variability. Here, we identify three axes by which fishers in Alaska,
one of the world's most productive fishing regions, can diversify fishing
activities. We show that, despite increasing specialization over the last 30
years, fishing permits with higher species diversity reduces individual revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of existing permits has a fishery-dependent effect on revenue and is
usually (93\% probability) associated with increased revenue uncertainty
the following year. Our results demonstrate that the most effective option for
individuals to decrease revenue variability is to purchase permits to
participate in additional or more diverse fisheries. However, this option is
expensive, often limited by regulations such as catch share programs, and
consequently unavailable to many individuals. With increasing environmental
variability in a changing climate, it will be particularly important that
individuals relying on natural resources for their livelihood have effective
strategies to reduce income variability.
 % (`r g2_dens`% probability)
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\input{ms}

\matmethods{
We obtained commercial fisheries landings and revenue data for all permit holders in Alaska from 1985 to 2014 from the Alaskan Commercial Fisheries Entry Commission (CFEC). These data represent a total of nearly 43 billion 2009 USD. We adjusted for inflation by setting all revenues to 2009 USD \cite{bea2016}. We implemented a number of filtering steps to focus on individuals actively engaged in commercial fishing (Supporting Information). In particular, we removed permit holders whose median revenue was less than \$10,000 USD to focus on individuals for whom commercial fishing formed a substantial part of their income.

We used combinations of state and federal fishing permits held by an individual in a specific year to define fishing ``strategies'', representing the species targeted (Supporting Information). For example, we combined permits that were otherwise the same except for vessel size (e.g.~a longliner fishing sablefish on a vessel under 60 ft was considered to have the same strategy as a longliner fishing sablefish on a vessel over 60 ft), and we combined a number of permits that were targeting a single species and differed only in region fished. For example, someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska Peninsula were both considered to have a ``herring roe'' permit group. For salmon permits, we formed salmon strategies based on similarities in gear type and species composition (Supporting Information).

After this process of combining permits, we were left with 23 unique fishing permit groups (Table~S\tabstrategies), which we aggregated within each person-year combination to form permit strategies. For example, if someone fished halibut and sablefish in the same year, their strategy for that year would be ``halibut-sablefish''. Specific details on our strategy definitions are available in the Supporting Information.

Our models focused on species diversity as a possible predictor of revenue and revenue variability while controlling for effort. For consistency with previous analyses {[}\cite{vedeld2007}; \cite{illukpitiya2008}; \cite{debela2012}; \cite{kasperski2013}; sethi2014{]}, we calculated effective species diversity for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in economics as the Herfindahl-Hirschman Index \cite{hirschman1964}) weighted by revenue \(R\):

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent where \(s\) indexes species from \(1\) through \(n\). We calculated effort or effective season length as the sum of the season lengths for each permit each year. Season length for a given permit was calculated as the number of days between the first instance of fishing a permit and the last instance of fishing a permit each year. For example, if someone fished for 30 days on a salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut permit, that individual's season length was calculated as 70 days.

\subsection{Hierarchical model}\label{hierarchical-model}

To jointly model revenue and revenue variability as a function of individual-level covariates, we extended the basic variance-function regression to a hierarchical Bayesian variance-function regression model with Gaussian errors predicting \(\log\) revenue, \(\log(R_{i,t})\), for each fisher \(i\) and time \(t\):

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
\end{equation}

\noindent Each year \(t\) of revenue for fisher \(i\) is assigned a strategy (permit set) indexed by \(j\) (Supporting Information, Table~\tabstrategies) and \(\mu_{i,t}\) is modelled as:

\begin{equation}
\label{eq:mean}
\mu_{i,t} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent where \(S_{i,t}\) is the log ratio of species diversity from year-to-year, \(\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)\), \(D_{i,t}\) is the log ratio of days fished from year-to-year, \(\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)\), and \(I_{i,t}\) is an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}\) (species diversity increasing; generalizing) and takes a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species diversity decreasing; specializing). Estimating a separate coefficient for the effect of increasing diversity and decreasing diversity allows the costs and benefits of these changes to be asymmetric (Fig.~S\figresid).

We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})\). We model change in revenue because preliminary analysis revealed non-stationary trends in revenue (resulting from experience or other unmodelled factors). This approach eliminates having to deal with the non-stationary time series and eliminates having to estimate thousands of mean-revenue intercepts associated with individual fishers. We removed person-year combinations without a subsequent year of revenue.

The \(\beta\) terms represent estimated coefficients, some of which are allowed to vary as ``random effect'' terms. The intercepts \(\beta_{0,j,t}\) are allowed to vary by strategy-year combinations, are centred on zero, and are constrained by a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent This approach, combined with the lack of a global intercept, constrains the mean change in log revenue across all fishers and years to be zero. I.e., we assume that the data are stationary after accounting for all covariates and including an intercept that varies with strategy-year combinations. Initial versions of the model confirmed that the global intercept along with strategy-level intercepts would be estimated at almost exactly zero. The estimated \(\beta_{j,t}\)'s represent years when people with a particular set of permits tended to collectively do better or worse than the long-term average (Fig.~S\figstryrs). Therefore, an individual's expected change in revenue (and variability) represents the expectation after accounting for the general trend in revenue for that set of permits.

The slopes \(\beta_{1,j}\) and \(\beta_{1,j}\) are then allowed to vary by strategy with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
\end{align}

\noindent The \(\beta_{1,j}\)'s and \(\beta_{2,j}\)'s represent how much a 1\% decrease and 1\% increase, respectively, in the ratio of species diversity from year-to-year within a strategy translates to a given percent change in the ratio of revenue from year-to-year. These are after controlling for changes in effort, \(\beta_3\), and the interaction between changes in effort and changes in species diversity, \(\beta_4\) and \(\beta_5\).

Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary systematically --- our variance-function regression model allows the residual variance to be heteroscedastic and vary as a function of predictors. Instead of estimating a single residual error scale term, \(\sigma\), we model the scale of the residual error with another hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\sigma_{i,t} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent Again, \(S_{i,t}\) and \(I_{i,t}\) represent, respectively, the log ratio of species diversity from year-to-year and an indicator variable for increasing or decreasing species diversity. We exponentiate the equation to ensure all scale parameters, \(\sigma_{i,t}\), are positive. Residuals from the mean component of our model clearly show a ``V''-shaped pattern, which the variance component of our model can capture (Fig.~S\figresid). Furthermore, the downside residuals (year-individual data points with lower than expected revenue) look qualitatively similar to the complete set of residuals (upside and downside) (Fig.~S\figresiddownside). Therefore, it seemed appropriate to model the variability as a proxy for revenue risk.

The \(\gamma\)'s are estimated coefficients, three of which are allowed to vary by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
\end{align}

\noindent The \(\gamma_{0,j}\) coefficients are modelled with two strategy-level predictors. The symbols \(M_j\) and \(E_j\) represent the mean species diversity and mean combined fishing season length (effort) for strategy \(j\). The coefficients \(\eta_1\) and \(\eta_2\) represent predictors in this strategy-level regression estimating effects across strategies.

We fit our models with Stan 2.14.1 \cite{standevelopmentteam2016, standevelopmentteam2016a} and R 3.3.2 \cite{rcoreteam2016}. Stan implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm \cite{hoffman2014} to perform Bayesian statistical inference. We assigned weakly informative priors on all parameters: \(N(0, 2^2)\) priors on all \(\beta\), \(\gamma\), and \(\eta\) parameters, and half-\(t(3, 0, 2)\) priors (i.e.~degrees of freedom of 3 and scale of 2) on all \(\sigma\) and \(\tau\). We ran four chains and 2,000 iterations and discarded the first 1,000 iterations of each chain as warm up. We checked for chain convergence visually with trace plots, ensured that \(\hat{R}<1.05\) (the potential scale reduction factor) and that the effective sample size, as calculated by the rstan R package \cite{standevelopmentteam2016}, was greater than 200 for all parameters \cite{gelman2014}. We verified that our model returned sensible estimates by comparing our estimates to an ad-hoc two-stage model fit with the lme4 R package \cite{bates2015}, where we first fit a mixed-effects regression model to the mean, and then modelled the absolute residuals from that first model in a second mixed-effects model representing the variance component. We also validated the model with simulated data, and were able to recover the true values (Supporting Information). Code used in our analysis is available at \url{https://github.com/NCEAS/pfx-commercial/portfolios/}.
}
\showmatmethods

\showacknow % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% If you see unexpected formatting errors, try commenting out this line
% as it can run into problems with floats and footnotes on the final page.
\pnasbreak

% Bibliography
\bibliography{refs}

\end{document}