### 2017-01-24 portfolio/text/anderson-etal-rev-pfx.tex
Incorporate edits

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Benefits and risks of diversification for individual fishers}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Sean C. Anderson}
\author[b]{Eric J. Ward}
\author[b]{Andrew O. Shelton}
\author[c]{Milo D. Adkison}
\author[c]{Anne H. Beaudreau}
\author[d]{Richard E. Brenner}
\author[e]{Alan C. Haynie}
\author[d]{Jennifer C. Shriver}
\author[f]{Jordan T. Watson}
\author[d]{Benjamin C. Williams}

\affil[a]{School of Aquatic and Fishery Sciences, Box 355020,
University of Washington, Seattle, WA 98195, USA}

\affil[b]{Conservation Biology Division, Northwest Fisheries Science Center,
National Marine Fisheries Service, National Oceanographic and
Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA}

\affil[c]{College of Fisheries and Ocean Sciences, University of Alaska Fairbanks,
17101 Point Lena Loop Rd, Juneau, AK, 99801, USA}

\affil[d]{Alaska Department of Fish and Game,
Division of Commercial Fisheries,
PO BOX 115526,
Juneau, AK, USA 99811-5526, USA}

\affil[e]{Alaska Fisheries Science Center,
National Marine Fisheries Service, National Oceanographic and
Atmospheric Administration, Bldg 4,
7600 Sand Point Way NE, Seattle, WA, 98115, USA}

\affil[f]{NOAA / National Marine Fisheries Service, Alaska Fisheries Science Center, Auke
Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA}

% Please give the surname of the lead author for the running footer
\leadauthor{Anderson}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{TODO}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
S.C.A, E.J.W, and A.O.S designed the study with contributions from all authors;
J.C.S. also provided data;
S.C.A. and E.J.W. analyzed the data and wrote the paper with input from all
authors.}
\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sandrsn@uw.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{
TODO $|$
TODO
}

\newcommand{\tabstrategies}{1}
\newcommand{\figfe}{1}
\newcommand{\figre}{2}
\newcommand{\figspokeall}{3}
\newcommand{\figresid}{4}
\newcommand{\figresiddownside}{5}
\newcommand{\figstryrs}{6}
\newcommand{\figifq}{7}
\newcommand{\figmirror}{8}
\newcommand{\fignoeffort}{9}

% 250:
\begin{abstract}
Individuals relying on the extraction of natural resources for their
livelihood---fishers, farmers, miners, and loggers, among others---face high
levels of income variability driven by a mix of environmental, biological,
management, and
economic factors. Key to managing these industries is identifying
how regulatory actions and individual behaviour affect income variability and, by
extension, economic stability and the sustainable use of natural resources. In
commercial fisheries, communities and vessels fishing a greater diversity of
species have less revenue variability than those fishing fewer species.
However, it is unclear if these benefits extend to the actions of individual
fishers and how year-to-year changes in diversification affect revenue and
revenue variability. Here, we evaluate two axes by which fishers in Alaska,
can diversify fishing
activities. We show that, despite increasing specialization over the last 30
years, fishing a set of permits with higher species diversity reduces individual revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of existing permits has a fishery-dependent effect on revenue and is
usually (93\% probability) associated with increased revenue uncertainty
the following year. Our results demonstrate that the most effective option for
individuals to decrease revenue variability is to purchase permits to
participate in additional or more diverse fisheries. However, this option is
expensive, often limited by regulations such as license limitation or catch share programs, and
consequently unavailable to many individuals. With increasing environmental
variability in a changing climate, it will be particularly important that
individuals relying on natural resources for their livelihood have effective
strategies to reduce income variability.
 % (`r g2_dens`% probability)
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\input{ms}

\matmethods{

We obtained commercial fisheries landings and revenue data for all permit
holders in Alaska from 1985 to 2014 from the Alaskan Commercial Fisheries Entry
Commission (CFEC). These data represent a total of nearly 43 billion 2009 USD.
We adjusted for inflation by setting all revenues to 2009 USD \cite{bea2016}.
We implemented a number of filtering steps to focus on individuals actively
engaged in commercial fishing (Supporting Information). In particular, we
removed permit holders whose median revenue was less than \$10,000 USD to focus
on individuals for whom commercial fishing formed a substantial part of their
income.

We used combinations of state and federal fishing permits held by an individual
in a specific year to define fishing ``strategies'', representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g.~a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished. For
example, someone fishing herring roe in Kodiak and someone fishing herring roe
in Alaska Peninsula were both considered to have a ``herring roe'' permit
group. For salmon permits, we formed salmon strategies based on similarities in
gear type and species composition (Supporting Information).

After this process of combining permits, we were left with 23 unique fishing
permit groups (Table~S\tabstrategies), which we aggregated within each
person-year combination to form permit strategies. For example, if someone
fished halibut and sablefish in the same year, their strategy for that year
would be ``halibut-sablefish''. Specific details on our strategy definitions
are available in the Supporting Information.

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses {[}\cite{vedeld2007}; \cite{illukpitiya2008}; \cite{debela2012};
\cite{kasperski2013}; sethi2014{]}, we calculated effective species diversity
for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's
diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in
economics as the Herfindahl-Hirschman Index \cite{hirschman1964}) weighted by
revenue \(R\):

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent where \(s\) indexes species from \(1\) through \(n\). We calculated
effort or effective season length as the sum of the season lengths for each
permit each year. Season length for a given permit was calculated as the number
of days between the first instance of fishing a permit and the last instance of
fishing a permit each year. For example, if someone fished for 30 days on a
salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
longline halibut permit, that individual's season length was calculated as 70
days.

\subsection{Hierarchical model}\label{hierarchical-model}

To jointly model revenue and revenue variability as a function of
individual-level covariates, we extended the basic variance-function regression
to a hierarchical Bayesian variance-function regression model with Gaussian
errors predicting \(\log\) revenue, \(\log(R_{i,t})\), for each fisher \(i\)
and time \(t\):

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
\end{equation}

\noindent Each year \(t\) of revenue for fisher \(i\) is assigned a strategy (permit set) indexed by \(j\) (Supporting Information, Table~\tabstrategies) and \(\mu_{i,t}\) is modelled as:

\begin{equation*}
  \begin{split}
\label{eq:mean}
\mu_{i,t} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -\\
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
  \end{split}
\end{equation*}

\noindent where \(S_{i,t}\) is the log ratio of species diversity from
year-to-year, \(\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)\),
\(D_{i,t}\) is the log ratio of days fished from year-to-year, \(\log
  \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)\), and \(I_{i,t}\) is
an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge
  \mathrm{div}_{i,t-1}\) (species diversity increasing; generalizing) and takes
a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species
diversity decreasing; specializing). Estimating a separate coefficient for the
effect of increasing diversity and decreasing diversity allows the costs and
benefits of these changes to be asymmetric (Fig.~S\figresid).

We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log
ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) =
  \log(R_{i,t}/R_{i,t-1})\). We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience
or other unmodelled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers. We removed
person-year combinations without a subsequent year of revenue.

The \(\beta\) terms represent estimated coefficients, some of which are allowed
to vary as ``random effect'' terms. The intercepts \(\beta_{0,j,t}\) are
allowed to vary by strategy-year combinations, are centred on zero, and are
constrained by a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent This approach, combined with the lack of a global intercept,
constrains the mean change in log revenue across all fishers and years to be
zero. I.e., we assume that the data are stationary after accounting for all
covariates and including an intercept that varies with strategy-year
combinations. Initial versions of the model confirmed that the global intercept
along with strategy-level intercepts would be estimated at almost exactly zero.
The estimated \(\beta_{j,t}\)'s represent years when people with a particular
set of permits tended to collectively do better or worse than the long-term
average (Fig.~S\figstryrs). Therefore, an individual's expected change in
revenue (and variability) represents the expectation after accounting for the
general trend in revenue for that set of permits.

The slopes \(\beta_{1,j}\) and \(\beta_{1,j}\) are then allowed to vary by
strategy with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
\end{align}

\noindent The \(\beta_{1,j}\)'s and \(\beta_{2,j}\)'s represent how much a 1\%
decrease and 1\% increase, respectively, in the ratio of species diversity from
year-to-year within a strategy translates to a given percent change in the
ratio of revenue from year-to-year. These are after controlling for changes in
effort, \(\beta_3\), and the interaction between changes in effort and changes
in species diversity, \(\beta_4\) and \(\beta_5\).

Whereas traditional hierarchical linear models assume homoscedasticity --- that
the residual variance does not vary systematically --- our variance-function
regression model allows the residual variance to be heteroscedastic and vary as
a function of predictors. Instead of estimating a single residual error scale
term, \(\sigma\), we model the scale of the residual error with another
hierarchical model with a similar form:

\begin{equation*}
\begin{split}
\label{eq:cv}
\sigma_{i,t} =
\exp (\gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -\\
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}).
\end{split}
\end{equation*}

\noindent Again, \(S_{i,t}\) and \(I_{i,t}\) represent, respectively, the log
ratio of species diversity from year-to-year and an indicator variable for
increasing or decreasing species diversity. We exponentiate the equation to
ensure all scale parameters, \(\sigma_{i,t}\), are positive. Residuals from the
mean component of our model clearly show a ``V''-shaped pattern, which the
variance component of our model can capture (Fig.~S\figresid). Furthermore, the
downside residuals (year-individual data points with lower than expected
revenue) look qualitatively similar to the complete set of residuals (upside
and downside) (Fig.~S\figresiddownside). Therefore, it seemed appropriate to
model the variability as a proxy for revenue risk.

The \(\gamma\)'s are estimated coefficients, three of which are allowed to vary
by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
\end{align}

\noindent The \(\gamma_{0,j}\) coefficients are modelled with two
strategy-level predictors. The symbols \(M_j\) and \(E_j\) represent the mean
species diversity and mean combined fishing season length (effort) for strategy
\(j\). The coefficients \(\eta_1\) and \(\eta_2\) represent predictors in this
strategy-level regression estimating effects across strategies.

We fit our models with Stan 2.14.1 \cite{standevelopmentteam2016,
  standevelopmentteam2016a} and R 3.3.2 \cite{rcoreteam2016}. Stan implements
the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm \cite{hoffman2014}
to perform Bayesian statistical inference. We assigned weakly informative
priors on all parameters: \(N(0, 2^2)\) priors on all \(\beta\), \(\gamma\),
and \(\eta\) parameters, and half-\(t(3, 0, 2)\) priors (i.e.~degrees of
freedom of 3 and scale of 2) on all \(\sigma\) and \(\tau\). We ran four chains
and 2,000 iterations and discarded the first 1,000 iterations of each chain as
warm up. We checked for chain convergence visually with trace plots, ensured
that \(\hat{R}<1.05\) (the potential scale reduction factor) and that the
effective sample size, as calculated by the rstan R package
\cite{standevelopmentteam2016}, was greater than 200 for all parameters
\cite{gelman2014}. We verified that our model returned sensible estimates by
comparing our estimates to an ad-hoc two-stage model fit with the lme4 R
package \cite{bates2015}, where we first fit a mixed-effects regression model
to the mean, and then modelled the absolute residuals from that first model in
a second mixed-effects model representing the variance component. We also
validated the model with simulated data, and were able to recover the true
values (Supporting Information). Code used in our analysis is available at
\url{https://github.com/NCEAS/pfx-commercial/portfolios/}. } \showmatmethods

\showacknow % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% If you see unexpected formatting errors, try commenting out this line
% as it can run into problems with floats and footnotes on the final page.
\pnasbreak

% Bibliography
\bibliography{refs}

\end{document}