### 2017-03-09 salmon/paper/salmon_paper_appendix_v3.Rmd
rearrange appendix

---
title: "Supplement for Ward et al., 'Reduced diversity and increased specialization of Alaska salmon fishermen' "
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
library(ggtern)
library(reshape2)
library(mgcv)
library(fpc)
library(gridExtra)
library(glmmTMB)
library(zoo)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
revenueThreshold = 10000

dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")

# single permit holders, salmon permits
dat.annual  = filter(dat.annual ,npermit==1 & substr(strategy_permit,1,1)=="S") %>%
  group_by(strategy_permit) %>% 
  mutate(nPeople = length(unique(p_holder))) %>% 
  filter(nPeople > 200) %>% 
  dplyr::select(-nPeople)
# drop troll
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S05B","S15B")==FALSE,]
# drop S04X, Y, Z because low participation (< 20) in many years
# S04W deleted because low median earnings
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S04W", "S04X","S04Y","S04Z")==FALSE,]

agg.annual = group_by(dat.annual, strategy_permit, year) %>% 
  summarize(people = length(unique(p_holder)),
    specDiv = mean(specDiv),
    meanrev = median(revenue),
    low = quantile(revenue,0.25),
    upp = quantile(revenue,0.75))

dat.diff = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]

agg.diff = group_by(dat.diff, strategy_permit) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
    cv = sd(log(revenue/revenue.prev)))

dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01M",
  "S03A","S03E"),]

# Also load in raw CFEC data for plots
#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
cfec$year = as.numeric(substr(cfec$landdate, 1, 4))

# only use people who are using just that permit - diversification within permit
cfec.raw = cfec
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]

# summarize annual data - total revenue and CV (3-year rolling)
annual_summary = group_by(cfec[which(cfec$year>0),], year, p_fshy) %>% 
  summarize(tot = sum(g_earn)) %>% arrange(p_fshy) %>% 
  group_by(p_fshy) %>% mutate(rollmean = c(NA,rollmean(tot,k=3),NA),
    rollsd = c(NA,rollapply(tot,width=3,sd),NA), rollcv = rollsd/rollmean)
#ggplot(annual_summary, aes(year, rollcv,color=p_fshy)) + geom_line()
#ggplot(annual_summary, aes(log(rollmean), rollcv, color=p_fshy)) + geom_point()

# calculate total earnings by person / permit / year
indiv_summary = group_by(cfec[which(cfec$year>0),], year, p_holder) %>% 
  mutate(n = length(unique(p_fshy))) %>% filter(n == 1) %>% 
  group_by(year, p_holder) %>% 
  summarize(g = sum(g_earn), p_fshy = p_fshy[1])
# add in missing data
people_year = expand.grid("year"=unique(cfec$year[cfec$year>0]), "p_holder"=unique(indiv_summary$p_holder))
indiv_summary = left_join(people_year, indiv_summary)
# apply rolling mean / sd to data
indiv_summary = group_by(indiv_summary, p_holder) %>% 
  mutate(rollmean = c(NA,rollapply(g,width=3,mean),NA),
    rollsd = c(NA,rollapply(g,width=3,sd),NA),
    rollcv = rollsd/rollmean)
# calculate averages across individuals over time
indiv_summary = group_by(indiv_summary, p_fshy, year) %>% 
  summarize(ind_mean = mean(rollmean,na.rm=T), 
    ind_cv = mean(rollcv,na.rm=T))

alldat = left_join(annual_summary, indiv_summary)
# this shows cvs that individuals experience are same as cvs of aggregate
#ggplot(alldat, aes(rollcv, ind_cv,color=p_fshy)) + geom_point()
#ggplot(alldat, aes(log(rollmean), ind_cv,color=p_fshy)) + geom_point()
#ggplot(alldat, aes(log(rollmean),log(ind_mean), color=p_fshy)) + geom_point(aes(alpha=year))
#ggplot(alldat, aes(year, log(rollmean))) + facet_wrap(~p_fshy) + geom_line()
#ggplot(alldat, aes(year, rollcv)) + facet_wrap(~p_fshy) + geom_line() + geom_line(aes(year, ind_cv, color="red"))
#ggplot(alldat, aes(year, log(ind_cv/rollcv), col=p_fshy)) + geom_line()
#ggplot(alldat, aes(log(rollcv), log(ind_cv),color=p_fshy)) + geom_point()

# Restrict analysis to people who keep same strategy - complicates days, etc
dat.salmon = dat.salmon[which(dat.salmon$strategy_permit==dat.salmon$strategy_permit.prev),]

dat.salmon$days.change = log((dat.salmon$days_permit+1) / (dat.salmon$days_permit.prev+1))
dat.salmon$year_permit = as.factor(paste(dat.salmon$year, dat.salmon$strategy_permit))

cfec = cfec[which(paste(cfec$p_holder,cfec$year) %in% paste(dat.salmon$p_holder,dat.salmon$year)),]

# Make come plots for the appendix of (1) median fishing dates over time, (2) length of fishing season over time (calculated as width of 95% intervals)
cfec.raw$month = as.numeric(substr(cfec.raw$day, 1,2))
cfec.raw$day = as.numeric(substr(cfec.raw$day, 4,5))
cfec.raw$jday = mdy.date(cfec.raw$month, cfec.raw$day, cfec.raw$year) - mdy.date(1, 1, cfec.raw$year)
# summarize the temporal distribution of the catches
cfec.raw = cfec.raw[which(is.na(cfec.raw$jday)==FALSE),]
seasons = group_by(cfec.raw, p_fshy, year) %>% 
  summarize(q = mean(sample(jday, size = 10000, replace=T, prob=g_earn)),
    lower = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.025),
    upper = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.975),
    length = upper - lower)

seasons = rename(seasons, permit = p_fshy)

seasons$permit[which(seasons$permit=="S 01A")] = "SE-seine"
seasons$permit[which(seasons$permit=="S 01E")] = "PWS-seine"
seasons$permit[which(seasons$permit=="S 01K")] = "KO-seine"
seasons$permit[which(seasons$permit=="S 01M")] = "AP-seine"
seasons$permit[which(seasons$permit=="S 03A")] = "SE-gillnet"
seasons$permit[which(seasons$permit=="S 03E")] = "PWS-gillnet"

appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

appendixPlot0_p2 = ggplot(seasons, aes(year, length, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Season length") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

\break

```{r makePlots, echo = FALSE}
# rename
agg.annual$label=NA
agg.annual$label[which(agg.annual$strategy_permit == "S01A")] = "SE-seine"
agg.annual$label[which(agg.annual$strategy_permit == "S01E")] = "PWS-seine"
agg.annual$label[which(agg.annual$strategy_permit == "S01H")] = "CI-seine"
agg.annual$label[which(agg.annual$strategy_permit == "S01K")] = "KO-seine"
agg.annual$label[which(agg.annual$strategy_permit == "S01L")] = "CH-seine"
agg.annual$label[which(agg.annual$strategy_permit == "S01M")] = "AP-seine"
agg.annual$label[which(agg.annual$strategy_permit == "S03A")] = "SE-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S03E")] = "PWS-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S03H")] = "CI-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S03M")] = "AP-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S03T")] = "BB-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S03W")] = "KU-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S04D")] = "YA-drift"
agg.annual$label[which(agg.annual$strategy_permit == "S04H")] = "CI-set"
agg.annual$label[which(agg.annual$strategy_permit == "S04K")] = "KO-set"
agg.annual$label[which(agg.annual$strategy_permit == "S04M")] = "AP-set"
agg.annual$label[which(agg.annual$strategy_permit == "S04T")] = "BB-set"

agg.annual = agg.annual[-which(agg.annual$label=="KU-drift"),]
appendixPlot1 = ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ label, scale="free") + ylab("Participation") + xlab("Year")

aggdiff = agg.diff[which(agg.diff$strategy_permit %in% unique(agg.annual$strategy_permit)),]
aggdiff$label = agg.annual$label[match(aggdiff$strategy_permit, agg.annual$strategy_permit)]

appendixPlot2_p1 = ggplot(aggdiff, aes(specDiv, log(meanrev), colour = cv, label=label)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

appendixPlot2_p2 = ggplot(aggdiff, aes(specDiv, cv, colour = log(meanrev), label=label)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

appendixPlot2_p3 = ggplot(aggdiff, aes(log(meanrev), cv, colour = specDiv, label=label)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

\break

# Appendix

## Overall trends in participation across permits

```{r figa1, fig.cap="Participation (unique people). Permits are coded by gear (purse seine, set and drift gillnets) and region ('AP' = Alaska Peninsula, 'BB' = Bristol Bay, 'CH' = Chignik, 'CI' = Cook Inlet, 'KO' = Kodiak, 'KU' = Kuskokwim, 'PWS' = Prince William Sound, 'SE' = Southeast, 'YA' = Yakutat). Effective species diversity is calculated as 1/Simpson's Diversity.", fig.pos="placeHere", echo=FALSE}
print(appendixPlot1)
```

\break

```{r figa2, fig.cap="Diversity, mean revenue and CV (aggregated across years). Permits are coded by gear (purse seine, set and drift gillnets) and region ('AP' = Alaska Peninsula, 'BB' = Bristol Bay, 'CH' = Chignik, 'CI' = Cook Inlet, 'KO' = Kodiak, 'KU' = Kuskokwim, 'PWS' = Prince William Sound, 'SE' = Southeast, 'YA' = Yakutat). Effective species diversity is calculated as 1/Simpson's Diversity.", fig.pos="placeHere", echo=FALSE}
gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
```

```{r, echo=FALSE}
#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
#saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
#cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))

# plot proportion of species landed by salmon permit
cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")

#cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
#cfec.raw = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]

g1 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01A" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("SE-seine") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name="Species",
    labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")

g2 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("PWS-seine") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g3 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01K" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("KO-seine") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g4 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01M" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("AP-seine") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name = "Species", labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")

g5 = group_by(cfec.raw[cfec.raw$p_fshy=="S 03A" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("SE-gillnet") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g6 = group_by(cfec.raw[cfec.raw$p_fshy=="S 03E" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec.raw$year > 0,], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("PWS-gillnet") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

dat.diff = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= 10000 & dat.diff$revenue.prev >= 10000), ]
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]

h1 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01A",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h2 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h3 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01K",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h4 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01M",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h5 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03A",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h6 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03E",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

```

```{r figa3, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit for seine fisheries in Southeast Alaska ('SE', S01A), Prince William Sound ('PWS', S01E) and Kodiak ('KO', S01K). Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye)", fig.height=3.2}
gridExtra::grid.arrange(g1, g2, g3, h1, h2, h3, nrow=2)
```

```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit, for seine fisheries on the Alaska Peninsula ('AP', S01M), and gillnet fisheries in Southeast Alaska ('SE', S03A) and Prince William Sound ('PWS', S03E).", fig.height=3.2} 
gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
```

\break

## Trends in revenue and prices

Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook. Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.  For most permit types, median revenue / person has generally increased, especially since early 2000s. For some permits, variability appears to have changed through time (quartile range getting smaller).

```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
# plot prices of species landed by salmon permit
cfec.raw$area = NA
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"

g1 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
  ylab("Revenue (millions) ") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
g1
```

\break

```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}

g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
p1 = group_by(g, spec, area) %>% 
  mutate(p = p/p[which(year==2002)]) %>% 
  group_by(year, area) %>%
  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change") + xlab("") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
p1

```

```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}

g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
# do log-diff of prices -- get price year before
g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
  paste(g$year+1, g$spec, g$area))]

p2 = group_by(g, spec, area) %>% 
  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") + xlab("") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
p2

g$logdiff = log(g$p/g$p_t1)
p3 = group_by(g, year, area) %>% 
summarize(sd = sd(logdiff,na.rm=TRUE)) %>% ggplot(aes(year, sd, color=area)) + geom_line() + xlab("Year") + ylab("SD of price changes across species") + ggtitle("Synchrony in price fluctuations across species") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

```

```{r, fig.pos="placeHere", fig.cap = "Standard deviation (across species, by area) for fluctuations in prices. Fluctuations are calculated as the log difference price data between sequential years. The standard deviation of log differenced data is interpreted as the coefficient of variation (CV)."}
p3
```

\break

## Causality between lagged abundance and prices

Previous work on long term trends in Alaska salmon markets have highlighted effects of changing production practices, including freezing (Knapp et al. 2012). Caches of frozen salmon from large harvest may have competitive effects on prices 1-2 years later. Using a Granger test for causality between first-differenced time series of harvest and prices, this effect is particularly true for coho and sockeye salmon, where abundance at time *t-1* or *t-2* improves predictions of prices at time *t*.

```{r, echo=FALSE}
library(knitr)
d = read.table("../../salmon/data-generated/granger_table.csv", sep=",", header=TRUE)
kable(d)
```

```{r echo=FALSE, eval=FALSE}
#Z = read.table("../../salmon/data-generated/granger_table_s01a.csv", header=TRUE)
```

\break

## Model Results

```{r figa5, fig.pos="placeHere", fig.cap="Time varying effects of increasing the effective species diversity of catch from 1 to 2 species, on log standard deviation of revenue.", echo=FALSE}

diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))
# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)
yearpermitold=yearpermit

# grab original data to look at permit names
diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold

est = rstan::extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev
x3 = log(specdiv/specdiv.prev) * specdiv.prev^2

df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$low = 0
df$high = 0

df$permits[which(df$permits=="S01A")] = "SE-seine"
df$permits[which(df$permits=="S01E")] = "PWS-seine"
df$permits[which(df$permits=="S01K")] = "KO-seine"
df$permits[which(df$permits=="S01M")] = "AP-seine"
df$permits[which(df$permits=="S03A")] = "SE-gillnet"
df$permits[which(df$permits=="S03E")] = "PWS-gillnet"

for(i in 1:nrow(df)) {
  #tmp = est$g0 + est$g0_strategy[,df$permit[i]] + 
  tmp = 
    x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]])# +
    #x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)
  df$pred[i] = median(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}

g1 = ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
  geom_line() +
  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit") + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
g1
```

```{r figa6, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit. These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
# grab original data to look at permit names
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")

years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))

# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)

yearpermitold=yearpermit
load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold
est = rstan::extract(mod, permute="TRUE")

yearpermit$b0 = apply(est$b0_str_yr,2,median)
yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)

yearpermit2 = yearpermit
yearpermit2$permit[which(yearpermit2$permit=="S01A")] = "SE-seine"
yearpermit2$permit[which(yearpermit2$permit=="S01E")] = "PWS-seine"
yearpermit2$permit[which(yearpermit2$permit=="S01K")] = "KO-seine"
yearpermit2$permit[which(yearpermit2$permit=="S01M")] = "AP-seine"
yearpermit2$permit[which(yearpermit2$permit=="S03A")] = "SE-gillnet"
yearpermit2$permit[which(yearpermit2$permit=="S03E")] = "PWS-gillnet"

g1 = ggplot(data=yearpermit2) + 
  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
  geom_line(aes(x=year,y=b0)) + 
  facet_wrap(~permit) + xlab("Year") + 
  ylab("Intercepts on mean") + 
  geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
g1
```

```{r figa7, eval=FALSE,echo=FALSE}
load("../../salmon/analysis/cv-model-simple.Rdata")
# groups here are 80% pink vs 97% pink

yrs = 1986:2014

hist(rstan::extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")

xx = exp(rstan::extract(mod)[["sd_yr"]])
df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")

offset = matrix(rstan::extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
xx = exp(extract(mod)[["sd_yr"]] + offset)
df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")

g1 = ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
g1
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Change in natural log of landed weight versus change in effective species diversity, by permit. Decreases in landed weight is often accompanied by reduced catch diversity."}
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
diffdat = filter(diffdat, !is.na(days.change))

g1 = ggplot(diffdat, aes(log(weight)-log(weight.prev), specDiv.change)) + geom_point(alpha=0.15, col="blue", size=0.4) + facet_wrap(~ strategy_permit) + xlab("Change in ln annual landed weight") + ylab("Change in effective species diversity") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
g1
```

```{r figa8, echo=FALSE, eval=FALSE}
# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E). This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
# grab original data to look at permit names
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold

est = extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev

df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$low = 0
df$high = 0
b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))

for(i in 1:nrow(df)) {
  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
  df$pred[i] = median(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}
df_s01e_b1 = df
#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
#  geom_line() +
#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")

```

```{r figa9, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns). Correlation between them is ~ 0.8. This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high. [EW side note: diversification may only be beneficial here when returns and prices are both low.]"}

yearpermit$b0 = apply(est$b0_str_yr,2,median)
yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)

S01E = yearpermit[yearpermit$permit=="S01E",]
names(S01E)[which(names(S01E)=="year")]="Year"

returns = readxl::read_excel("../../salmon/paper/salmon_returns_PWS.xlsx")
pinkPWS = returns[,c("Year", "Pink_total", "Chum_total")]
pinkPWS = pinkPWS[which(pinkPWS$Year >= 1975),]
pinkPWS$diff = c(NA, diff(log(pinkPWS$Pink_total)))

pinkPWS = dplyr::left_join(pinkPWS, S01E)

g1 = ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
g1
```

\break

### Season length

There appears to be no significant trend in season length (defined as the difference between 95% CIs) or mean julian day for any fishery. In all cases, the dates are weighted by the earnings associated with landed catch.

```{r, fig.cap="Trends in mean julian day and season length by fishery", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
gridExtra::grid.arrange(appendixPlot0_p1, appendixPlot0_p2, ncol = 1)
```