### 2017-05-29 portfolio/text/strategy-table.tex
Update strategy table compilation

% latex table generated in R 3.3.2 by xtable 1.8-2 package
Label & Description & Mean spp. diversity & Main species \\ 
  \midrule
Dun C & Dungeness crab & 1.0 & crab \\ 
  King C & King crab & 1.0 & crab \\ 
  Herr - pound & Herring, pound & 1.0 & herring \\ 
  Sea cuc & Sea cucumber & 1.0 & sea cucumber \\ 
  Tan C & Tanner crab, pot gear & 1.0 & crab \\ 
  Hal - ll & Halibut, longline & 1.0 & halibut \\ 
  Fin - pot & Finfish, pot gear, statewide & 1.0 & cod \\ 
  Herr roe - gill & Herring roe, gillnet & 1.0 & herring \\ 
  Sab & Sablefish, longline, statewide & 1.0 & sablefish \\ 
  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB & 1.1 & sock \\ 
  Fin - ll & Finfish, longline, statewide & 1.1 & pollock \\ 
  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB & 1.2 & sock \\ 
  S - seine - CH & Salmon, purse seine, Chignik & 1.3 & sock \\ 
  S - seine - PWS & Salmon, purse seine, PWS & 1.4 & pink, chum \\ 
  Fin - otter & Finfish, otter trawl, statewide & 1.5 & cod \\ 
  S - set - YAK & Salmon, set gillnet, Yakutat & 1.7 & coho, sock \\ 
  S - troll & Salmon, power troll, statewide & 1.8 & chnk, coho \\ 
  S - seine - CI, K & Salmon, purse seine, Cook Inlet and Kodiak & 1.9 & pink, sock \\ 
  S - drift - PWS & Salmon, drift gillnet, PWS & 2.0 & coho, pink, sock, chum \\ 
  S - seine - SE & Salmon, purse seine, Southeast Alaska & 2.1 & pink, sock, chum \\ 
  S - seine - AP & Salmon, purse seine, Alaska Penninsula & 2.2 & pink, sock, chum \\ 
  S - set - NS/KU & Salmon, set gillnet, Norton Sound and Kuskokwim & 2.5 & chnk, coho, sock, chum \\ 
  S - drift - SE & Salmon, drift gillnet, Southeast Alaska & 2.5 & coho, pink, sock, chum \\ 
   \bottomrule