### 2017-07-07 portfolio/text/ms.Rmd
Edit last paragraph of introduction; dealing with reviewer comment

```{r set-knitr-options, echo=FALSE}
library("knitr")
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results="hide")
```
```{r get-dat}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

<!-- # Introduction -->

\dropcap{I}t can be difficult for individuals to sustain a livelihood from
natural resource extraction. These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to income variability from economic sources,
such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].
<!-- Could add anthropogenic perturbations (e.g., Exxon Valdez) to this
list or in the next paragraph. -->

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may diversify their crops or
include off-farm income sources to buffer against environmental and market
shocks, as well as longterm climatic trends and seasonality [@ellis2000;
@barrett2001; @ellis2004]. However, otherwise well-intentioned regulations may
limit how individuals diversify, or incentivize against diversification. For
instance, crop subsidies in the U.S. may incentivize some farms to specialize
(e.g. dairy, grains, soybeans; @mishra2002). In other instances, market forces
(e.g. prices or demand) may discourage the adoption of diverse strategies and
help promote specialization [@bradshaw2004].

Diversification may benefit commercial fishers, given the extreme economic risk
that some individuals experience, particularly relative to terrestrial
agriculture [@sethi2012b]. The majority of risk insurance programs available to
terrestrial farmers (e.g. crop insurance programs) are not available to
commercial fishers [@kasperski2013]. However, fishers may be able to lower risk
by diversifying their fishing portfolio. They can, for example, target as many
species as allowed by a permit, own multiple permits, or diversify spatially and
fish in multiple regions. The effectiveness of these diversification strategies
for increasing revenue and reducing revenue variability remains uncertain. Theory and empirical studies demonstrate that there can be substantial
benefits to specialization for both farmers and fishers [@smith1986;
@abdulai2000; @arnason2005]. Furthermore, the majority of research examining the
effects of diversification on revenue do so at aggregate levels that combine
individuals (e.g., counties or communities; [@kasperski2013;
@sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship 
among diversity, fishing revenue, and fishing revenue variability at the level 
of the individual.

<!-- important to acknowledge that IFQ regs have been implemented partly for economic -->
<!-- reasons and have been successful by some economic measures. Might even want to -->
<!-- insert the above sentence. -->

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. Fishers must own a permit to participate in virtually
all fisheries [@hilborn2003a]. While such regulations have helped reduce
overfishing and in many fisheries have increased profits, they limit the
available strategies to reduce income variability. For example, many permits
limit fishing to a single species, diversifying by purchasing additional fishing
permits can be expensive[^1], and some fisheries limit the entry of new
participants through individual fishing quotas (IFQs) [@grimm2012].

[^1]: For example, it cost ~\$192K USD for a king and tanner crab permit in
Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.

<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth
a footnote to address the potential critique that at the indivdual level we
don’t observe their wages in non-fishing activities.  I’m not sure if this is
worth bringing up, but this is always the case anyway with vessels where the
skipper may do other things. triue, and for others - KH and Cl et al. ...-->

Using a unique dataset collected over 30 years on individual fishers, rather
than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals, it is
possible to track people that switch fishery permits over time or use multiple
vessels and differentiate fishers that share the same vessel. We develop
a hierarchical variance-function regression model [@davidian1987; @hall1989] to
simultaneously understand the effect of species
diversity within fisheries catches on individual revenue and revenue variability.
Our model builds on results from vessel [@kasperski2013], fleet [@sethi2012b],
and community-level [@sethi2014; @cline2017] variability,
and although we model revenue *variability*, we herein equate revenue
variability with risk for consistency with Kasperski and Holland [@kasperski2013].
We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions).
We use the model to evaluate two axes by which fishers in Alaska can
diversify fishing activities:
(1) choice of permit strategy (i.e., permit or set of permits fished), and (2)
choice of what species to fish *within* a permit strategy. To our knowledge,
this is the first application of this type of model in any natural resource
field.

\section*{Results}

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold a single permit, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders---compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple-permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of the catch that generated this revenue declined somewhat from
1985--2000 but has since remained steady (Fig. \ref{fig:ts}D). This trend
appears largely driven by an increase in salmon specialists in the late 1980s and the
1990s, disappearing if salmon license holders are excluded (Fig. \ref{fig:ts}D).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.45\textwidth]{fig1-jk.pdf}

\caption{Trends in revenue and diversification over time. (A) Map of major fishing
areas in Alaska, USA. (B) Total revenue in billions of USD from `r
range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
holders. (C) Proportion of single permit holders for all fisheries (thick black
line) and with permit holders omitted each year who fished any of the `r
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.45\textwidth]{stan-across-strategy-variability2.pdf}

\caption{Individuals who fish permits and permit combinations (permit
strategies) with higher species diversity have lower expected revenue
variability. On the y-axis, dots, thick lines, and thin lines represent
posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
common permit strategies. Thick black line, and dark and light grey shaded
regions indicate median, 50\%, and 95\% credible intervals of strategy-level
regression built into the hierarchical model. Estimated variability represents
expected standard deviation for a permit holder who does not change species
diversity or days fished from year-to-year. }

\label{fig:across}
\end{center}
\end{figure}
<!-- TODO AB: To me it is more intuitive that the darker, hotter colors would
represent higher values. Consdier reversing the color scale to go from red
(highest # of permit holders) to yellow (lowest #).  -->

Individual revenue variability decreases exponentially with higher mean fished
species diversity across combinations of state and federal fishing permits
("strategies"; Fig. \ref{fig:across}). This pattern is
qualitatively similar with or without individuals targeting salmon. Controlling for effort, strategies with
one unit higher mean species diversity had on average `r h1[[2]]`% (`r
h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. As examples,
revenue variability was highest for the herring roe gillnet and king crab fisheries.
Most strategies with the lowest estimated revenue variability combined halibut
with a salmon permit and had high levels of fished species diversity (Fig.
\ref{fig:across}).

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in year-to-year revenue
variability from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\figspokeall). For example, individuals fishing a halibut permit plus one
additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=0.65\textwidth]{stan-gg-spoke2.pdf}

\caption{Diversifying permits reduces risk and usually increases revenue. Two
examples of expected revenue variability and median revenue for owners of (A)
individual halibut and (B) gillnet herring roe permits (dark blue) compared with
owners of the same permit plus one more (medium blue) or two more (light
blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviations of year-to-year revenue rates
of return. Angled lines connects permit sets that differ by one permit.}

\label{fig:spokes}
\end{center}
\end{figure*}

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given strategy (set of permits) is more
variable (Fig. \ref{fig:within}). After accounting for fishing effort and
changes in annual average revenue for fishers with the same set of permits,
specializing tends to be associated with increased revenue but also increased
variability (Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\figmirror).
On average, there is a ~0.5% increase in the
year-to-year revenue ratio per 1% decrease in species diversity, and across
strategies weighted equally, `r g2_dens`% of the probability density supports
a positive relationship between specialization and revenue. However, when
not controlling for effort, the average effect of specializing on revenue
becomes more neutral (Fig. S\fignoeffort).

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.46\textwidth]{stan-offset-break-anti-spaghetti-specializing.pdf}

\caption{Effect of specializing on year-to-year changes in revenue and revenue
variability for individual permit holders \emph{within} the `r nstr` most common
strategies. Changing species diversity from year-to-year is associated with
higher variability for individuals within the strategies above the zero line.
Specializing is associated with greater expected revenue ratios to the right of
the zero line and lower expected revenue ratios to the left of the zero line.
For example, individuals with setnet permits for salmon in Prince William Sound,
Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
their revenue variability (vertical axis) for a 1\% decrease in their ratio of
year-to-year species diversity. These effects are estimated holding fishing
season length constant from year-to-year. Dots and line segments represent
medians and 50\% credible intervals of posteriors. 
In other words, the axes represent coefficient values from the strategy-level slopes. Moving along an axis (say from left to right on the horizontal axis) represents an increase in how steep the slope is between specialization and the year-to-year revenue ratio.
The effect of generalizing
(increasing species diversity) is similar (Fig. S\figmirror).}

\label{fig:within}
\end{center}
\end{figure}

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a strategy tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre, S\figmirror).
This effect was strongest for individuals with a halibut and sablefish permit.
Individuals with this strategy experienced a ~2% increase in variability for
a 1% change in the year-to-year ratio of species diversity (Figs
\ref{fig:within}, S\figmirror). 
<!-- Becoming more generalized in species diversity -->
<!-- also leads to increased revenue variability (Fig. S\figmirror A), suggesting -->
<!-- that the act of changing the number of species fished for a given set of permits -->
<!-- is a contributor to increased variability of revenue. -->

<!-- TODO mention the quadrants in figure 4? -->

Alaskan halibut and sablefish fisheries were converted to individual fishery 
quota (IFQ) management in 1995, which potentially confounds our estimates with a
major regulatory change. Therefore, we refit our models allowing all 
strategy-level effects involving halibut and sablefish permits to vary 
separately before and after the introduction of IFQs. The introduction of IFQs 
was associated with a reduction in revenue variability of ~20--30% for
individuals with only halibut or sablefish permits (Fig. S\figifq A;
although other changes in fish populations occurred concurrently).
Furthermore, for individuals within most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\figifq B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\figifq B).

\section*{Discussion}

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
@edwards2004]. However, in practice, the ability of individuals to effectively
diversify is limited by regulation and market forces. Here we document the
effect of individual-level changes in fished species diversity on revenue and
revenue variability across two scales. We find that individuals with sets
of permits that allow high levels of species diversity tend to have lower
revenue variability and adding a fisheries permit usually increases revenue and
decreases revenue variability. However, within the constraints of a permit strategy
and accounting for effort, specializing tends to increase revenue while changes to fished species diversity tend to correlate with greater revenue variability,
although there are considerable differences amongst fisheries. Thus, the most effective
mechanism by which species diversity can lower revenue variability for
individual fishers---switching or adding fishing permits---is expensive,
limited by regulations, and consequently unavailable to many individuals.

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
findings within strategies, with respect to individual year-to-year behavior,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. 
Similar efficiency gains from specialization are well known in 
agricultural fields [@ikerd1990; @abdulai2000].
Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990].
Third, diversifying---a lack of
specialization---can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].
Because specialization inherently holds risk from ``putting all one's eggs in one
basket,'' it is not suprising that a decline in the most valuable resource
would lead to a shift towards diversification.
<!-- Response to A1 -->
Conversely, a specialization may appear as a response to increased revenue.
A fisher who achieves high revenue from a hotspot of a particular fish species
one year may choose to stop or pause fishing thereby indirectly limiting catch
diversity. For example, pink and chum salmon are the dominant species caught in
purse seine fisheries in Prince William Sound, with chum salmon generally
harvested earlier in the season than pink salmon. If fishers targeting chum salmon
achieve high revenue early in the season, they may choose to fish less later in
the season.

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, local knowledge and experience are
a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
fishing patterns may sometimes be associated with a learning or adaptation phase
as individuals become accustomed to new locations, gear, or species.
<!-- For example, many individuals fishing herring in Alaska in the 1970s --> 
<!-- AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies. Otherwise, statements like this one aren’t really anchored to anything concrete.  For example, what does increased specialization look like in the Hal ll permit, which showed one of the greatest effects of specialization on revenue diversity? How many permits could a Hal ll permit holder potentially fish? This goes back to my comment re: Table 1 (see p. 15) 
sea cuc - H W book - learning...
herring - alaska 1970s - learn to fish for roe ... [ref]
salmon - hatchery production - change target from diverse to most abundant... PWS seine - pink salmon or SE
Hal ll + S troll...
-->
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers instead chose to target other species or landed
more non-target species,
adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behavior [@hilborn1985; @hilborn1985a;
@branch2006] and therefore not in our model, that might
make a specific individual's revenue less predictable if species diversity
changes.

<!-- there are highly variable costs among fisheries with gear and permits in particular 
but it is not clear that costs vary systematically with the diversity of the catch 
there are opportunity costs to ... -->
<!-- gear costs incur considerable capital investment and some maintenance costs 
Permit types and gear types are closely linked and therefore fishers largely rely on the same gear regardless of whether you fish a specialized or diverse catch 
people have limited flexibility to modify gear within permits which are largely determined by regulations 
some permit combinations would allow a fisher to fish multiple permits with a single boat while others would require multiple boats
just expand our list of cost factors and call it a day 
costs such as leasing vessels from other fishers
maintaining equipment  -->

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial expense. Related barriers are apparent in
other natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008].
<!-- Response to reviewer comment A2 -->
While we were unable to incorporate costs, it is possible that 
including license and gear purchase costs along with others such as costs of
gear maintenance, catch transportation, and selling to a processor
would alter the relationship between catch diversification and profits.
<!-- unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here. -->
Another major barrier to
diversification can be management restrictions. In commercial fisheries,
regulations that have been shown to reduce overfishing---including improved
assessments [@melnychuk2017] and individual quota systems 
[@costello2008]---may ultimately restrict individual diversification [@smith1986; @kasperski2013]
or at least make developing a diverse fishing portfolio more expensive. However,
individual quota systems, in particular, may also reduce income variability
(Fig. S\figifq A) perhaps offsetting the effects of reduced diversification
[@holland2016].

Our analysis provides a synthetic view of the effects of species diversity on 
individuals over 30 years in one of the most productive fishing regions in the 
world. 
<!-- Response to reviewer comment B2 -->
Although we focus on Alaska, and expect our results are largely generalizable to other productive and intensely managed fishing regions, it is possible that the benefits of diversification would be stronger in regions where fish stocks are highly variable or in regions where permit costs are lower and more freely available.
<!--Overall, the large temporal and geographical scope of our analysis potentially obscures 
time or regional-variation in the consequences of diversification on revenue and
revenue variability. -->
Furthermore, in addressing the overall pattern, we have
averaged out the potential signal of several major discrete events, such as the
*Exxon Valdez* oil spill in 1989, that was a significant disruption to the
management, ecological, and social structure of fisheries in Alaska
[@paine1996]. In addition to revenue or income variability, it is important to
understand the benefits of diversification in the context of rare and extreme
disturbances which occur in natural systems worldwide [@anderson2017].

With increasing environmental variability and extreme events in a changing
climate [@easterling2000], it is particularly important that individuals who
rely on natural resources as a primary source of income, such as fishers,
farmers, and forestry workers, have effective strategies to reduce risk. While
diversification can provide a powerful form of risk buffering, we have shown
that the relationship between individual changes to diversification and income
variability is complex and context specific. Furthermore, external market forces
can interact with management regulations, which have other considerable
benefits, to restrict individual diversification opportunities and promote
specialization. 
From a management perspective, metrics such as interannual variability in individual fishing revenue or fraction of single permit holders within communities could serve as bioeconomic indicators.
Although variability in individual revenue is not yet actively
incorporated into management decisions, analyses such as ours help to explain
the range of diversification strategies available, trade-offs inherent in such
management decisions, and how these trade-offs may change over time.

<!-- vim: textwidth=80
-->